---
title: "Cangulatus"
author: "Andy"
date: "2024-09-13"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      eval = TRUE,
                      cache = TRUE)
```

```{r}
rm(list = ls())
```


#Library
```{r}
library(ggplot2)
library(tidyr) #for %>%
library(dplyr) #for select function
library(ggridges)
```

```{r}

data_folder <- "/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/"
```

#Genome Survey
```{r}

kmer_types <- c("17", "19", "21", "23")

all_data <- data.frame()

for (k in kmer_types) {
  file_path <- paste0(data_folder, '00-Genome_Survey/', k, 'mer.histo')
  tmp <- read.table(file_path)
  tmp <- subset(tmp, V1 >= 5 & V1 <= 500)
  tmp$Kmer <- paste0("k = ", k)
  all_data <- rbind(all_data, tmp)
}
p <- ggplot(all_data, aes(x = V1, y = V2, color = Kmer)) + 
  geom_line(size = 0.8) +
  xlab("K-mer depth") + 
  ylab("Number of distinct k-mers") +
  coord_cartesian(ylim = c(0, 5e+06)) +
  theme_bw() + # õĮ┐ńö© theme_bw ÕÅ»õ╗źÕ┐½ķĆ¤ÕŠŚÕł░Õ╣▓ÕćĆńÜäĶāīµÖ»
  theme(
    legend.position = c(0.7, 0.6), # Õ░åÕøŠõŠŗń¦╗ÕŖ©Õł░ÕÉłķĆéõĮŹńĮ«
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    panel.background = element_blank()
  )
p
write_excel(all_data, file = paste0(data_folder, "00-Genome_Survey/kmer_data.xlsx"), row.names = FALSE)
```


#Genome Evolution
##1A)ks
```{r}
ks_data = read.table(paste0(data_folder,"/04-Evolution_Analysis/02-Synteny_Analysis/01-KaKs/merge_ks.txt"),header = T)

#define a function that identify density plot peak
densFindPeak <- function(x) {
  td <- density(x,adjust = 1/2)
  maxDens <- which.max(td$y)  #the position of the element which the maximal value in a vector.
  list(x=td$x[maxDens],y=td$y[maxDens])
}

CanA_ks <- ks_data %>% filter(Species == 'CanA_CanA') %>% pull(Ks)

CanA_ks_limit1 <- CanA_ks[CanA_ks >= 0 & CanA_ks <= 1]
CanA_ks_peak1 = densFindPeak(CanA_ks_limit1)
CanA_ks_peak1 #0.1981212

CanA_ks_limit2 <- CanA_ks[CanA_ks >= 1 & CanA_ks <=2.5]
CanA_ks_peak2 = densFindPeak(CanA_ks_limit2) #2.078022
CanA_ks_peak2 #2.078022

#Twi_Twi peak
Twi_ks <- ks_data %>%  filter(species == 'Twi_Twi') %>% select(ks) 
Twi_ks_limit1 <- Twi_ks$ks[Twi_ks$ks >= 0 & Twi_ks$ks <=1]
Twi_ks_peak1 = densFindPeak(Twi_ks_limit1) #0.1595298


#Rco_Rco peak
Rco_ks <- ks_data %>%  filter(species == 'Rco_Rco') %>% select(ks) 
Rco_ks_limit1 <-Rco_ks$ks[Rco_ks$ks >= 0 & Rco_ks$ks <=1]
Rco_ks_peak1 = densFindPeak(Rco_ks_limit1) #0.8350758

#Pti_Pti peak
Pti_ks <- ks_data %>%  filter(species == 'Pti_Pti') %>% select(ks) 
Pti_ks_limit1 <-Pti_ks$ks[Pti_ks$ks >= 0 & Pti_ks$ks <=1]
Pti_ks_peak1 = densFindPeak(Pti_ks_limit1) #0.2299008

#Vvi_Vvi peak
Vvi_ks <- ks_data %>%  filter(species == 'Vvi_Vvi') %>% select(ks) 
Vvi_ks_limit1 <-Vvi_ks$ks[Vvi_ks$ks >= 1.5 & Vvi_ks$ks <=2.5]
Vvi_ks_peak1 = densFindPeak(Vvi_ks_limit1) #1.560698

#CanA_Vvi peak
CanA_Vvi_ks <- ks_data %>%  filter(species == 'CanA_Vvi') %>% select(ks) 
CanA_Vvi_ks_limit1 <-CanA_Vvi_ks$ks[CanA_Vvi_ks$ks >= 1 & CanA_Vvi_ks$ks <=2]
CanA_Vvi_ks_peak1 = densFindPeak(CanA_Vvi_ks_limit1) #1.38778

#CanA_Twi
CanA_Twi_ks <- ks_data %>%  filter(species == 'CanA_Twi') %>% select(ks) 
CanA_Twi_ks_limit1 <- CanA_Twi_ks$ks[CanA_Twi_ks$ks >= 0 & CanA_Twi_ks$ks <=0.6]
CanA_Twi_ks_peak1 = densFindPeak(CanA_Twi_ks_limit1) #0.07694491

#CanA_Pti
CanA_Pti_ks <- ks_data %>%  filter(species == 'CanA_Pti') %>% select(ks) 
CanA_Pti_ks_limit1 <- CanA_Pti_ks$ks[CanA_Pti_ks$ks >= 0 & CanA_Pti_ks$ks <=0.6]
CanA_Pti_ks_peak1 = densFindPeak(CanA_Pti_ks_limit1) #0.5819391

#CanA_Rco
CanA_Rco_ks <- ks_data %>%  filter(species == 'CanA_Rco') %>% select(ks) 
CanA_Rco_ks_limit1 <- CanA_Rco_ks$ks[CanA_Rco_ks$ks >= 0 & CanA_Rco_ks$ks <=0.6]
CanA_Rco_ks_peak1 = densFindPeak(CanA_Rco_ks_limit1) #0.552019

######
# 1. µé©ńÜäÕćĮµĢ░õ┐ØµīüõĖŹÕÅś’╝łµŚóńäČµé©ÕĆŠÕÉæõ║ÄĶ┐ÖõĖ¬ńēłµ£¼’╝ē
densFindPeak <- function(x) {
  td <- density(x, adjust = 1/2)
  maxDens <- which.max(td$y)
  list(x=td$x[maxDens], y=td$y[maxDens])
}

# 2. Loop
species_list <- c('CanA_CanA','Twi_Twi','Rco_Rco','Pti_Pti','Vvi_Vvi',
                  'CanA_Twi','CanA_Pti','CanA_Rco','CanA_Vvi')

peaks_list <- lapply(species_list, function(sp) {
  sub_ks <- ks_data %>% filter(Species == sp) %>% pull(Ks)
  
  if (sp == 'CanA_CanA') {
    # µĀĖÕ┐āõ┐«µŁŻ’╝ÜõĮ┐ńö© $x µÅÉÕÅ¢ Ks ÕĆ╝
    p1 <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 1])$x
    p2 <- densFindPeak(sub_ks[sub_ks >= 1 & sub_ks <= 2.5])$x
    # µŁżµŚČ c(p1, p2) µś»õĖĆõĖ¬ÕīģÕÉ½õĖżõĖ¬µĢ░ÕŁŚńÜäÕÉæķćÅ’╝īõ╝Üńö¤µłÉ 2 ĶĪīµĢ░µŹ«
    return(data.frame(Species = sp, peak = c(p1, p2)))
  } else if (sp == 'Vvi_Vvi') {
    p <- densFindPeak(sub_ks[sub_ks >= 1.5 & sub_ks <= 2.5])$x
    return(data.frame(Species = sp, peak = p))
  } else if (sp == 'CanA_Twi') {
    p <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 0.6])$x
    return(data.frame(Species = sp, peak = p))
  } else {
    p <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 1.5])$x
    return(data.frame(Species = sp, peak = p))
  }
})

# 3. ńÄ░Õ£©µē¦ĶĪīÕÉłÕ╣Č’╝īµēĆµ£ēµĢ░µŹ«µĪåķāĮń╗¤õĖĆõĖ║ 2 ÕłŚ’╝łSpecies, peak’╝ē
peaks_df <- do.call(rbind, peaks_list)

# µēōÕŹ░ń╗ōµ×£µ¤źń£ŗ
print(peaks_df)

####
## Ridges plot Õ▒▒ĶäŖÕøŠ
#define level of species
ks_data$species <- factor(ks_data$species,levels = c('CanA_CanA','Twi_Twi','Rco_Rco',
                                                     'Pti_Pti','Vvi_Vvi','CanA_Twi',
                                                     'CanA_Pti','CanA_Rco','CanA_Vvi'))

ks_data %>% ggplot(aes(x=ks,y=species,fill = species)) + 
  geom_density_ridges(scale = 3, bandwidth = 0.08, color = F) +
  geom_vline(xintercept = 0.1981212, color = 'black', linetype = 'dashed') + 
  #geom_vline(xintercept = 0.6235556, color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.1595298,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.8350758,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.2299008,color = 'black', linetype = 'dashed') +
  geom_vline(xintercept = 2.078022,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 1.38778,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.07694491,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.5819391,color = 'black', linetype = 'dashed') +
  #geom_vline(xintercept = 0.552019,color = 'black', linetype = 'dashed') +
  scale_x_continuous(limits =c(0,2.5),expand = c(0.015,0)) +
  scale_y_discrete(expand = c(0.05,0)) + 
  scale_fill_manual(values = c("#DC143C",'#B0C4DE', "#6495ED", "#FF8C00", '#2E8B57', "#FF6347", "#87CEFA", "#F4A460", '#3CB371')) +
  guides(fill = FALSE) +
  theme_classic() + 
  theme_ridges(font_size = 13, grid = FALSE) +
  theme(axis.title.y = element_blank()) #axis.title delete

```

##1B)Ks
```{r}
ks_data = read.table(paste0(data_folder,"/04-Evolution_Analysis/02-Synteny_Analysis/01-KaKs/merge_ks.txt"),header = T)

densFindPeak <- function(x, bw = 0.5) {
  x <- x[!is.na(x)] # µÄÆķÖżń╝║Õż▒ÕĆ╝
  if(length(x) < 2) return(NA)
  # õĮ┐ńö©Õ»åÕ║”ÕćĮµĢ░Ķ«Īń«ŚÕłåÕĖāµø▓ń║┐
  td <- density(x, bw = bw)
  # µēŠÕł░ y ĶĮ┤µ£ĆÕż¦ÕĆ╝Õ»╣Õ║öńÜä x ĶĮ┤ÕØÉµĀć
  maxDens <- which.max(td$y)
  return(td$x[maxDens])
}

species_list <- c('CanA_CanA','Twi_Twi','Rco_Rco','Pti_Pti','Vvi_Vvi',
                  'CanA_Twi','CanA_Pti','CanA_Rco','CanA_Vvi')

# 4. µē╣ķćÅĶ«Īń«ŚÕ│░ÕĆ╝
peaks_list <- lapply(species_list, function(sp) {
  # µ│©µäÅ’╝ÜĶ┐ÖķćīõĮ┐ńö©õ║å Species ÕÆī Ks (õĖÄ header = T ÕÉÄńÜäÕłŚÕÉŹõĖĆĶć┤)
  sub_ks <- ks_data %>% filter(Species == sp) %>% pull(Ks)
  
  # µĀ╣µŹ«ńē®ń¦Źµ╝öÕī¢ńē╣ÕŠüĶ«ŠÕ«ÜĶ┐ćµ╗żÕī║ķŚ┤
  if (sp == 'CanA_CanA') {
    # µ¤źµēŠĶ┐æµ£¤ WGD Õ│░ (0-1) ÕÆīÕÅżĶĆüõĖēÕĆŹÕī¢Õ│░ (1-2.5)
    p1 <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 1])
    p2 <- densFindPeak(sub_ks[sub_ks >= 1 & sub_ks <= 2.5])
    return(data.frame(species = sp, peak = c(p1, p2)))
  } else if (sp == 'Vvi_Vvi') {
    # ĶæĪĶÉäķĆÜÕĖĖÕÅ¬õ┐ØńĢÖÕÅżĶĆüńÜä gamma õ║ŗõ╗ČÕ│░
    p <- densFindPeak(sub_ks[sub_ks >= 1.5 & sub_ks <= 2.5])
    return(data.frame(species = sp, peak = p))
  } else if (sp == 'CanA_Twi') {
    # ķÆłÕ»╣ÕłåÕī¢ķØ×ÕĖĖĶ┐æµ£¤ńÜäńē®ń¦Ź’╝īń╝®ń¬äÕī║ķŚ┤õ╗źń▓ŠÕćåÕ«ÜõĮŹ
    p <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 0.6])
    return(data.frame(species = sp, peak = p))
  } else {
    # ÕģČõ╗¢ńē®ń¦ŹńÜäÕłåÕī¢µł¢ WGD Õ│░ķĆÜÕĖĖÕ£© 0-1.5 õ╣ŗķŚ┤
    p <- densFindPeak(sub_ks[sub_ks >= 0 & sub_ks <= 1.5])
    return(data.frame(species = sp, peak = p))
  }
})

# ÕÉłÕ╣Čń╗ōµ×£õĖ║ń╗śÕøŠµēĆķ£ĆńÜäµĢ░µŹ«µĪå
peaks_df <- do.call(rbind, peaks_list)

# µēōÕŹ░ń╗ōµ×£µ¤źń£ŗ
print(peaks_df)

peaks_df <- data.frame(
  Species = c('CanA_CanA', 'CanA_CanA'),
  peak = c(0.2247, 2.0891),
  label = c("Recent WGT", "Gamma event")
)

peaks_df <- data.frame(
  Species = c('CanA_CanA', 'CanA_CanA'),
  peak = c(0.2131514, 1.8716905),
  label = c("Recent WGT", "Gamma event")
)
# 3. Ķ«ŠńĮ«ńē®ń¦ŹķĪ║Õ║Å
#species_order <- c('CanA_Vvi','CanA_Rco', 'CanA_Pti','CanA_Twi','Rco_Rco', 'Pti_Pti', 'Vvi_Vvi','Twi_Twi','CanA_CanA')
species_order <- c('CanA_CanA', 'Twi_Twi', 'Rco_Rco', 'Pti_Pti', 'Vvi_Vvi', 
                   'CanA_Twi', 'CanA_Pti', 'CanA_Rco', 'CanA_Vvi')

ks_data$Species <- factor(ks_data$Species, levels = species_order)
peaks_df$Species <- factor(peaks_df$Species, levels =species_order)

# 4. ń╗śÕøŠ
ggplot(ks_data, aes(x = Ks, y = Species, fill = Species)) +
  geom_density_ridges(scale = 2.5, bandwidth = 0.08, color=NA,alpha = 0.9) +
  
  # Ķć¬ÕŖ©µśĀÕ░ä CanA_CanA ńÜäĶÖÜń║┐
  geom_vline(data = peaks_df, aes(xintercept = peak), 
             color = "black", linetype = "dashed", size = 0.5) +
  
  # õ┐«µŁŻ’╝Üy µśĀÕ░äńÄ░Õ£©ÕÅ»õ╗źµēŠÕł░ 'Species' õ║å
  geom_text(data = peaks_df, aes(x = peak, y = Species, label = label), 
            vjust = -1, hjust = -0.1, size = 3.5, fontface = "bold.italic") +

  scale_x_continuous(limits = c(0, 2.5), expand = c(0.01, 0), name = "Ks value") +
  scale_fill_manual(values = c("red",'purple', "#6495ED", "#FF8C00", '#2E8B57', 
                               "#FF6347", "#87CEFA", "#F4A460", '#3CB371')) +
  theme_ridges(font_size = 12, grid = FALSE) +
  theme_classic() +
  theme(
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = "none"
  )

#####################
species_order <- c('CanA_CanA', 'Twi_Twi', 'Rco_Rco', 'Pti_Pti', 'Vvi_Vvi', 
                   'CanA_Twi', 'CanA_Pti', 'CanA_Rco', 'CanA_Vvi')

# Õ░å species ĶĮ¼µŹóõĖ║ factor Õ╣ČÕÅŹĶĮ¼ķĪ║Õ║Å’╝łÕøĀõĖ║ ridges Ķć¬õĖŗĶĆīõĖŖÕĀåÕÅĀ’╝ē
ks_data$Species <- factor(ks_data$Species, levels = rev(species_order))
peaks_df$species <- factor(peaks_df$species, levels = rev(species_order))

# 4. Õ╝ĆÕ¦ŗń╗śÕøŠ
ggplot(ks_data, aes(x = Ks, y = Species, fill = Species)) +
  # ń╗śÕłČÕ▒▒ĶäŖń║┐’╝ībandwidth=0.08 ńĪ«õ┐ØÕ╣│µ╗æÕ║”õĖÄÕ│░ÕĆ╝Ķ«Īń«ŚõĖĆĶć┤
  geom_density_ridges(scale = 3, bandwidth = 0.08, color = "white", alpha = 0.8) +
  
  # Ķć¬ÕŖ©µśĀÕ░äĶ«Īń«ŚÕć║ńÜäÕ│░ÕĆ╝ĶÖÜń║┐
  geom_vline(data = peaks_df, aes(xintercept = peak), 
             color = "black", linetype = "dashed", size = 0.4) +
  
  # Ķ«ŠńĮ«ÕØÉµĀćĶĮ┤ĶīāÕø┤õĖÄµ»öõŠŗ
  scale_x_continuous(limits = c(0, 2.5), expand = c(0.01, 0), name = "Ks value") +
  
  # Ķć¬Õ«Üõ╣ēķó£Ķē▓ķģŹĶē▓µ¢╣µĪł
  scale_fill_manual(values = c("#3CB371", "#F4A460", "#87CEFA", "#FF6347", 
                               "#2E8B57", "#FF8C00", "#6495ED", "#B0C4DE", "#DC143C")) +
  
  # Ķ░āµĢ┤õĖ╗ķóśń╗åĶŖé
  theme_ridges(font_size = 12, grid = FALSE) +
  theme_classic() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(vjust = 0),
    legend.position = "none"
  )

# ÕłøÕ╗║ń╗¤õĖĆńÜäÕćĮµĢ░µØźĶÄĘÕÅ¢KsÕ│░ÕĆ╝
get_ks_peaks <- function(species_name, ks_data, ks_ranges = list(c(0, 1))) {
  result <- list()
  
  # ńŁøķĆēńē®ń¦ŹµĢ░µŹ«
  species_ks <- ks_data %>% 
    filter(species == species_name) %>% 
    pull(ks)
  
  if(length(species_ks) == 0) {
    warning(paste("No data found for species:", species_name))
    return(NULL)
  }
  
  # õĖ║µ»ÅõĖ¬KsĶīāÕø┤Ķ«Īń«ŚÕ│░ÕĆ╝
  for(i in seq_along(ks_ranges)) {
    range <- ks_ranges[[i]]
    ks_subset <- species_ks[species_ks >= range[1] & species_ks <= range[2]]
    
    if(length(ks_subset) > 1) {
      peak <- densFindPeak(ks_subset)
      result[[paste0("peak", i)]] <- peak$x
    } else {
      result[[paste0("peak", i)]] <- NA
    }
  }
  
  return(result)
}

# Ķ«ŠńĮ«ńē®ń¦ŹÕÆīÕ»╣Õ║öńÜäKsĶīāÕø┤
species_config <- list(
  'CanA_CanA' = list(ranges = list(c(0, 1), c(1, 2.5))),
  'Twi_Twi' = list(ranges = list(c(0, 1))),
  'Rco_Rco' = list(ranges = list(c(0, 1))),
  'Pti_Pti' = list(ranges = list(c(0, 1))),
  'Vvi_Vvi' = list(ranges = list(c(1.5, 2.5))),
  'CanA_Vvi' = list(ranges = list(c(1, 2))),
  'CanA_Twi' = list(ranges = list(c(0, 0.6))),
  'CanA_Pti' = list(ranges = list(c(0, 0.6))),
  'CanA_Rco' = list(ranges = list(c(0, 0.6)))
)

# µē╣ķćÅĶ«Īń«ŚµēĆµ£ēÕ│░ÕĆ╝Õ╣ČÕŁśÕé©
all_peaks <- list()

for(sp in names(species_config)) {
  peaks <- get_ks_peaks(sp, ks_data, species_config[[sp]]$ranges)
  all_peaks[[sp]] <- peaks
}

# ÕłøÕ╗║õĖĆõĖ¬µĢ░µŹ«µĪåÕŁśÕé©µēĆµ£ēÕ│░ÕĆ╝’╝łńö©õ║ÄĶć¬ÕŖ©ń╗śÕłČń½¢ń║┐’╝ē
peak_df <- data.frame()
for(sp in names(all_peaks)) {
  for(peak_name in names(all_peaks[[sp]])) {
    peak_value <- all_peaks[[sp]][[peak_name]]
    if(!is.na(peak_value)) {
      peak_df <- rbind(peak_df, data.frame(
        species = sp,
        peak_name = peak_name,
        peak_value = peak_value
      ))
    }
  }
}

# µēōÕŹ░Õ│░ÕĆ╝ń╗ōµ×£
print("KsÕ│░ÕĆ╝Ķ«Īń«Śń╗ōµ×£’╝Ü")
for(sp in names(all_peaks)) {
  cat(paste("\n", sp, ":\n"))
  print(unlist(all_peaks[[sp]]))
}

# Ķ«ŠńĮ«ńē®ń¦ŹķĪ║Õ║Å
ks_data$species <- factor(ks_data$species, 
                         levels = c('CanA_CanA', 'Twi_Twi', 'Rco_Rco',
                                    'Pti_Pti', 'Vvi_Vvi', 'CanA_Twi',
                                    'CanA_Pti', 'CanA_Rco', 'CanA_Vvi'))

# ķó£Ķē▓Ķ«ŠńĮ«
species_colors <- c(
  "CanA_CanA" = "#DC143C",
  "Twi_Twi" = 'purple',
  "Rco_Rco" = "#6495ED",
  "Pti_Pti" = "#FF8C00",
  "Vvi_Vvi" = '#2E8B57',
  "CanA_Twi" = "#FF6347",
  "CanA_Pti" = "#87CEFA",
  "CanA_Rco" = "#F4A460",
  "CanA_Vvi" = '#3CB371'
)

# ń╗śÕłČÕ▒▒ĶäŖÕøŠ - Ķć¬ÕŖ©µĘ╗ÕŖĀÕ│░ÕĆ╝ń½¢ń║┐
p <- ks_data %>% 
  ggplot(aes(x = ks, y = species, fill = species)) + 
  geom_density_ridges(scale = 3, bandwidth = 0.08, color = F, alpha = 0.8) +
  scale_x_continuous(limits = c(0, 2.5), expand = c(0.015, 0)) +
  scale_y_discrete(expand = c(0.05, 0)) + 
  scale_fill_manual(values = species_colors) +
  guides(fill = "none") +
  theme_classic() +
  theme_ridges(font_size = 13, grid = FALSE) +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  xlab("Ks (Synonymous Substitution Rate)")

# Ķć¬ÕŖ©µĘ╗ÕŖĀµēĆµ£ēÕ│░ÕĆ╝ń½¢ń║┐
for(i in 1:nrow(peak_df)) {
  p <- p + geom_vline(xintercept = peak_df$peak_value[i], 
                     color = "black", 
                     linetype = "dashed",
                     alpha = 0.5,
                     size = 0.5)
}

# µśŠńż║ÕøŠĶĪ©
print(p)

# ÕÅ»ķĆē’╝Üõ┐ØÕŁśÕøŠĶĪ©
# ggsave("ks_distribution_plot.pdf", p, width = 10, height = 8)
# ggsave("ks_distribution_plot.png", p, width = 10, height = 8, dpi = 300)
```

#Stacked Bar
```{r}
rm(list = ls())
setwd("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/04-Phylogenetic_Analysis/01-OG")
#BiocManager::install("ggpubr")
library(ggplot2)
library("ggpubr")
dat <- read.table('Species_Statistics.txt', head = T, sep ='\t')
dat
#define the order of 
type <- factor(dat$feature,levels = c('Unclustered-genes','Unique paralogs','Multiple-copy-orthologs','Single-copy-orthologs'))
type
p <- ggplot(dat, aes(x=species, y=counts, fill=type)) +
     geom_bar(stat = "identity", width = 0.6, position = "stack")+
     scale_fill_manual(values =c("orange","#00FF00","red","blue"))+
     labs(x="",y="")+
     coord_cartesian(ylim = c(0,60000)) + 
    scale_y_continuous(breaks = c(0,60000,5000))+
     theme(axis.text.x = element_text(angle = 70, hjust = 1), #XĶĮ┤µ¢ćµ£¼Ķ«ŠńĮ«
           axis.ticks.length.y = unit(0.15,'cm'), 
           axis.ticks.length.x = unit(0.15,'cm'), #’Ż®
           legend.position = c(0.8,0.8),  #show legend
          panel.grid.major = element_blank(),panel.grid.minor = element_blank(),#
          axis.line = element_line(size = 1, colour = "black"), #
          text = element_text(size = 14, family = "serif") ,#
          panel.background = element_blank()) + 
 #   scale_y_continuous(breaks = seq(0,120000,10000),expand = c(0,0))+ #????Y???╠Č╚Ż?╚ź????X????ŽČ
    scale_y_continuous(expand = c(0,0)) #ÕØÉµĀćÕ╗Čõ╝Ė’╝īńĪ«õ┐ØÕøŠÕĮóÕģāń┤ĀĶ”åńø¢Ķć│ÕØÉµĀćĶĮ┤
p       
ggsave('./stacked_bar.pdf',p,width = 10,height = 8)        #)+ 
#+ geom_text(aes(label = counts),position = position_stack(vjust = 0.5),size =5) #????????═╝?─▒?
  #theme_bw()  

#====================================================================================
p2 <- ggplot(dat, aes(x=species, y=counts, fill=feature)) +
  geom_bar(stat = "identity", width = 0.6, position = "stack")+
  scale_fill_manual(values =c("#00FF00","orange","red","blue","#800080"))+
  labs(x=NULL,y=NULL,fill = NULL)+
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+ 
  coord_cartesian(ylim = c(70000,125000)) + #
  scale_y_continuous(breaks = c(70000,120000,10000))+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(),#
        axis.line = element_line(size = 1, colour = "black")) #
        #text = element_text(size = 14, family = "serif") )#
p2 

ggarrange(p2,p1,heights = c(1/2, 1/2),ncol = 1, nrow =2,common.legend = TRUE, align = "v")

```


##2)Gene Family
```{r}
# BiocManager::install("ggtree")
# BiocManager::install("ggnewscale")
# BiocManager::install("ggtreeExtra")
library(ggtree)
library(ggnewscale)
library(ggtreeExtra)

RNAexpr <- read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/01-GeneFam/02-TPS/01-Phylogenetic_Tree/CanA_fpkm.csv')

tps_lst = read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/01-GeneFam/02-TPS/01-Phylogenetic_Tree/tps_genes_expression.csv',header = T,sep=',')

idx = which(RNAexpr$Gene_ID %in% tps_lst$ID)
RNAexpr <- RNAexpr[idx,]

RNAexpr$Flower1 <- rowMeans(RNAexpr[ , c(2, 3)], na.rm=TRUE)
RNAexpr$Flower2 <- rowMeans(RNAexpr[ , c(4, 5)], na.rm=TRUE)
RNAexpr$Flower3 <- rowMeans(RNAexpr[ , c(6, 7)], na.rm=TRUE)
#RNAexpr$Flower <- rowMeans(RNAexpr[ , 2:7], na.rm=TRUE)

RNAexpr$Leaf1 <- rowMeans(RNAexpr[ , c(8, 9)], na.rm=TRUE)
RNAexpr$Leaf2 <- rowMeans(RNAexpr[ , c(10, 11)], na.rm=TRUE)
RNAexpr$Leaf3 <- rowMeans(RNAexpr[ , c(13, 14)], na.rm=TRUE)
#RNAexpr$Leaf <- rowMeans(RNAexpr[ , c(8:11,13,14)], na.rm=TRUE)

RNAexpr$Root1 <- rowMeans(RNAexpr[ , c(15, 16)], na.rm=TRUE)
RNAexpr$Root2 <- RNAexpr[ , 17]
RNAexpr$Root3 <- rowMeans(RNAexpr[ , c(18, 19)], na.rm=TRUE)
#RNAexpr$Root <- rowMeans(RNAexpr[ , 15:19], na.rm=TRUE)

#RNAexpr$Stem <- rowMeans(RNAexpr[ , c(20, 21)], na.rm=TRUE)

RNAexpr = RNAexpr[,c(-2:-19)]

rownames(RNAexpr) <- RNAexpr$Gene_ID
expr_dat <- log2(RNAexpr[,-1]+1)

CaTPS_tree <- read.tree("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/01-GeneFam/02-TPS/01-Phylogenetic_Tree/RAxML_bestTree.CaATPS-Tree")

p<- ggtree(CaTPS_tree,layout = "circular",size=0.1,branch.length = "none")+
  geom_tiplab(align = T,size=2)

p2 <-gheatmap(p,
              expr_dat,
              offset = 6,
              width = 1.5,
              colnames_angle = 95,
              legend_title = "log2(FPKM)",
              high="#D32F2F",low = "#FFF7F3",
              font.size = 2)
p2
```


#Transcript_Analysis

##KEGG_Enrich
```{r eval=FALSE}
rm(list=ls())
#setwd("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_analysis/05-Duplication/")
setwd('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/04-Phylogenetic_Analysis/02-Expansion')
library(ggplot2)

goinput<-read.csv("CaA.expanded.significant.genes_KEGG.csv",sep= ",", header = T)
goinput$Rich.factor=goinput$intersection_size/goinput$term_size

x = goinput$Rich.factor
#reorder() function to reorder the term_name according to the values in each Rich.factor
y = reorder(factor(goinput$term_name,levels = goinput$term_name),
                    goinput$Rich.factor)   

p = ggplot(goinput, aes(x, y)) + geom_point(aes(size=intersection_size,color=adjusted_p_value))+  
  scale_color_gradient(low = "red", high = "blue") +   #creates a two color gradient (low-high)
  scale_radius(range = c(10,20))+  #set radius of circles  
  guides(color= guide_colorbar(order = 2),size=guide_legend(order=1))+
  labs(color=expression("P value"), 
       size="Gene Number",
       x="Rich factor",
       y="KEGG pathway",
       #title=module
  ) + 
  theme_bw() +
  theme(
  legend.title = element_text(size=20)) +        
  theme(legend.text = element_text(size=15)) +   
  theme(legend.key.size = unit(60, "pt")         
        #  #legend.key.height = unit(35, "pt"),
        #legend.key.width = unit(55, "pt")
        )+
  theme(axis.text.x = element_text(colour = "black",
                                   size = 20, vjust =1 ),      #x????ūó??????ąĪ
         axis.text.y = element_text(colour = "black",
                                   size = 20, hjust =1 ),      #y????ūó??????ąĪ
        axis.title.x = element_text(margin=margin(10, 5, 0, 0),
                                  color = "black",size = 30),  #xy??????ąĪ
        axis.title.y = element_text(angle=90))   


ggsave(p,filename = "KEGG.pdf", width = 16,height = 16,dpi = 300)
```

##PCA
```{r eval=FALSE}
#install.packages('devtools')
library(devtools)
# options(repos = c(
#     fawda123 = 'https://fawda123.r-universe.dev',
#     CRAN = 'https://cloud.r-project.org'))
# 
# # Install ggord
# install.packages('ggord')
library(ggord)
library(ggplot2)
# Enable the r-universe repo
#=========================µ¢╣µ│Ģ1=====================================================================
FPKM_data = read.csv(file='/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/01-PCA/CanA_fpkm.csv',header = T,sep = ',',check.names = F,comment.char = "",row.names = 1) #ńē╣µ«Ŗń¼”ÕÅĘ
#TPM_data = read.csv(file = 'CaA_TPM_matrix.csv',header = T, sep = ',',check.names = F,comment.char = "",row.names = 1)
#c_data = FPKM_data[ ,1:20]
c_data = FPKM_data[ ,1:20]
raw_data = t(as.matrix(c_data)) #Õ░åĶĪ©ĶŠŠķćÅµĆ╗ĶĪ©ÕÅŹĶĮ¼’╝īPCAÕćĮµĢ░Ķ«Īń«ŚńÜäµś»Õ▒Ģńż║ńÜäń¼¼õĖĆÕłŚID
class(raw_data)  #matrix
str(raw_data)

pca_group <- factor(c(rep('Flower_1',2),rep('Flower_2',2),rep('Flower_3',2),rep('Leaf_1',2),rep('Leaf_2',2),rep('Leaf_3',1),rep('Leaf_4',2),rep('Root_1',2),rep('Root_2',1),rep('Root_3',2),rep('Stem_2',1),rep('Stem_2',1)))
pac_group1 <- factor(c(rep('Flower',6),rep('Leaf',7),rep('Root',5),rep('Stem',2)))#Õłåń╗ä

pca_data = prcomp(raw_data) #prcompµś»RĶć¬ÕĖ”ńÜäpcaÕłåµ×ÉÕćĮµĢ░
pcam.sum = summary(pca_data)
p <- ggord(pca_data,
           grp_in = pac_group1,
           arrow=0,
           vec_ext=0,
           txt =NULL,
           ellipse_pro=0.85,
           vec_lab=TRUE)+
  theme_bw() 
#Õ£©ggordÕÅéµĢ░õĖŁÕŖĀpoly = FALSEµżŁÕ£åµŚĀĶāīµÖ»;ellipse=FµŚĀµżŁÕ£å;ellipse_pro=ńĮ«õ┐ĪÕī║ķŚ┤’╝īÕå│Õ«ÜµżŁÕ£åÕøŖµŗ¼ńÜäÕż¦Õ░Åhull=TµżŁÕ£åÕåģµ£ēĶÅ▒ÕĮó
#obslab=TRUEÕŖĀµĀĘÕōüÕÉŹ
p
?ggord
##########################µ¢╣µ│Ģ2######################################
pca.sum_1<-as.data.frame(pcam.sum$x)
pca.sum_1$group_1 <-c("Flower_1-1","Flower_1-2","Flower_2-1","Flower_2-2","Flower_3-1","Flower_3-2","Leaf_1-1","Leaf_1-2","Leaf_2-1","Leaf_2-2","Leaf_3-1","Leaf_4-1","Leaf_4-2","Root_1-1","Root_1-2","Root_2-1","Root_3-1","Root_3-2","Stem_1","Stem_2")
ggplot(pca.sum_1, aes(x=PC1,y=PC2,colour=group_1))+geom_point(size=4)+theme_bw() 
write.table(pca_data,"datra",quote = F, sep = "\t")
pca_data = data.frame()
head(pca_data)
probesetvar = apply(pca_data,1,var)
write.table(probesetvar,"probesetvar",quote = F, sep = "\t")
head(probesetvar)
ord <- prcomp(pca_data[, 1:6]) 
head (ord)l
p <- ggord(ord,pca_data$transcript_id, cols = )
p
qplot(PCA1,PCA2,data=ord,colour=factor(type))+theme_classic()
plot(pca_data$PCA1,pca_data$PCA2,type="p",col=factor(pca_data$type))

```


#Differential Gene Expression Analysis
###DEseq2
```{r}
############################## STEP 1:µĢ░µŹ«ńÜäĶĮĮÕģźÕÆīÕćåÕżć ##########################################
#BiocManager::install("DESeq2")
library(DESeq2)

countData <- read.csv("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/02-DESeq/A-gene_count.csv",sep=",",header = T) #gene expression datafile

countData$Flower1 <- rowMeans(countData[ , c(2, 3)], na.rm=TRUE)
countData$Flower2 <- rowMeans(countData[ , c(4, 5)], na.rm=TRUE)
countData$Flower3 <- rowMeans(countData[ , c(6, 7)], na.rm=TRUE)

countData$Leaf1 <- rowMeans(countData[ , c(8, 9)], na.rm=TRUE)
countData$Leaf2 <- rowMeans(countData[ , c(10, 11)], na.rm=TRUE)
countData$Leaf3 <- rowMeans(countData[ , c(13, 14)], na.rm=TRUE)

countData$Root1 <- rowMeans(countData[ , c(15, 16)], na.rm=TRUE)
countData$Root2 <- countData[ , 17]
countData$Root3 <- rowMeans(countData[ , c(18, 19)], na.rm=TRUE)

countData = countData[,c(-2:-19)]

rownames(countData)<-countData[,1]

countData<-countData[,-1]

samplename <- factor(c(rep("Stem",2),
                       rep("Flower",3),
                       rep("Leaf",3),
                       rep("Root",3)
                       ),
                     levels = c("Stem",
                                "Flower",
                                "Leaf",
                                "Root"
                                )
                     )

#batch <- factor(c(rep()))#Ķ«ŠńĮ«ÕÄ╗µē╣µ¼ĪµĢłÕ║öńÜäÕÅśķćÅ
colData <- data.frame(row.names = colnames(countData),samplename) #µĘ╗ÕŖĀÕÄ╗µē╣µ¼ĪµĢłÕ║öńÜäÕÅśķćÅbatch
colData
head(countData)
samplename

#Create DESeqDataSet Object#µ¤źń£ŗÕÄ╗µē╣µ¼ĪµĢłÕ║öńÜäÕÅśķćÅ
dds <- DESeqDataSetFromMatrix(round(countData), 
                              colData, 
                              design = ~ samplename)  #µĘ╗ÕŖĀÕÄ╗µē╣µ¼ĪµĢłÕ║öńÜäÕÅśķćÅbatch
keep <- rowSums(counts(dds) >= 10) >= 2  
#the gene has counts >=10 in 2 or more samples;
#reduce noise or false positive
#reduce the memory size of the dds data objective
#increase the speed of count modeling with DESeq2
#improve visualization
#keep <- rowSums(counts(dds)) >= 10       
dds <- dds[keep, ]


##############################STEP 2:õĖ╗µłÉÕłåÕłåµ×É#########################################
#Remove batch effect using vst transformation (rlog for small sample sizes <30; vst for samples >30) 
vsdata <- vst(dds, blind=FALSE) 
#Remove batch effect;'batch' is a design variable
assay(vsdata) <- limma::removeBatchEffect(assay(vsdata), vsdata$batch) 

p <-plotPCA(vsdata, intgroup = "samplename",returnData = T)
p

pcaData <-plotPCA(vsdata, intgroup=c("samplename"),returnData = T)
pcaData <- pcaData[order(pcaData$samplename,decreasing=F),]
table(pcaData$samplename)
palette()
colors()
#Plot the PCA data (PC1 vs. PC2), with custom colors for each sample group
plot(pcaData[,1:2],
     pch = 19,
     col= c(rep("gray",2), #stem samples
            rep("red",3), # Flower 
            rep("green3",3), # Leaf
            rep("cyan",3)) #root
     )
                  
text(pcaData[,1],pcaData[,2],row.names(pcaData),cex = c(0.5,0.5), font = 1)

#µĀ╣µŹ«PCAplot.pdfÕÆīreturnData ÕÄ╗ķÖżńö¤ńē®ÕŁ”ķćŹÕżŹõ╣ŗķŚ┤ĶĘØń”╗ĶŠāĶ┐£ńÜäµĀĘµ£¼######################

#############  STEP 3: Differential Gene Expression Analysis ###############################

#dds_norm <- DESeq(dds)
dds_norm <- DESeq(dds, minReplicatesForReplace = Inf) #normalized counts without  replacing any potential outlier valuesõĖŹÕēöķÖżoutliers; õĖÄcookscutoffõĮ£ńö©ńøĖÕÉī.Ķ»źÕÅéµĢ░ÕĮōń”╗ńŠżÕĆ╝ÕŹĀµ»öĶŠāÕż¦’╝īµĢ░ķćÅĶŠāÕżÜµŚČõĮ┐ńö©’╝īÕģĘõĮōĶ¦üÕŹ░Ķ▒Īń¼öĶ«░ŃĆé
res <- results(dds_norm)
res
############################ STEP 3.1:Õ║ōÕģŗĶĘØń”╗Õłżµ¢Łń”╗ńŠżÕĆ╝ ######################################
p <- boxplot(log10(assays(dds_norm)[["cooks"]]), range=0, las=2,outline=TRUE)
p
dev.off()

####################    STEP 4:ÕłøÕ╗║Õ╣ČÕłćµŹóÕł░Ķ”üµ»öĶŠāµĀĘµ£¼ńÜäÕĘźõĮ£ńø«ÕĮĢ         ############################
comparisons <- combn(levels(colData$samplename), 2, simplify = FALSE)
results_list <- list()

for (i in seq_along(comparisons)) {
  tissue1 <- comparisons[[i]][1]
  tissue2 <- comparisons[[i]][2]
  res <- results(dds_norm, contrast = c("samplename", tissue2, tissue1))
  #dds_norm$comparison_group <- factor(dds_norm$samplename,levels = c(tissue1,tissue2))
  #res <- results(dds_norm, contrast = c("comparison_group", tissue2, tissue1))
  res_filtered <- res[which(abs(res$log2FoldChange) > 2 & res$padj < 0.05), ]
  
  comparison_name <- paste(tissue1, "vs", tissue2, sep = "_")
  results_list[[comparison_name]] <- res_filtered
  cat("Comparison done:", tissue1, "vs", tissue2, "\n")
}

degs_lists <- lapply(results_list, function(res) {
  # Extract gene identifiers where the adjusted p-value < 0.05
  rownames(res[which(res$padj < 0.05), ])
})

all_degs <- unlist(degs_lists)

length(all_degs)
unique_degs <- unique(all_degs)
length(unique_degs)


summary(res)
  
#########################################################################################################
#########################                                              ##################################
#########################           STEP 5:ńŁøķĆēÕĘ«Õ╝éĶĪ©ĶŠŠÕ¤║ÕøĀ            ##################################
#########################                                              ##################################
#########################################################################################################

#resOrdered <- res[order(res$pvalue), ] #µÄÆÕ║Å,õĮ┐ńö©ÕÉÄ’╝īÕÉÄķØóµĢ░µŹ«õĮ┐ńö©resOrderedŃĆé
aa=sum(res$padj<0.05 & res$log2FoldChange  > 1.5 , na.rm = TRUE)
bb=sum(res$padj<0.05 & res$log2FoldChange  < -1.5, na.rm = TRUE)
cc=sum(res$padj<0.05 & abs(res$log2FoldChange)  > 1.5, na.rm = TRUE)
dd=as.data.frame(c(aa,bb,cc))
colnames(dd)=c(paste(treatmentgroup,"-",controlgroup))
rownames(dd)=c("up","down","total")
write.table(dd,quote = F,row.names = T,append = T,file = "../Statistics.txt",col.names = T)
#NormalizeData=as.data.frame(counts(dds_norm,normalize=TRUE)) #µēōÕŹ░µēĆµ£ēµĀćÕćåÕī¢ńÜäcountsµĢ░µŹ«
NormalizeData=subset(as.data.frame(counts(dds_norm,normalize=TRUE)),
                     select=c(rownames(subset(colData(dds_norm),colData(dds_norm)$condition==treatmentgroup | colData(dds_norm)$condition==controlgroup))))#ÕÅ¬µēōÕŹ░µ»öÕ»╣µĀĘµ£¼ńÜäµĀćÕćåÕī¢µĢ░µŹ«
res_data <- merge(as.data.frame(res),#res or resOrdered
                  NormalizeData,
                  by="row.names",
                  sort=FALSE)
res_data=subset(res_data, padj!= "NA"&pvalue!= "NA",)
up_DEG <- subset(res_data, padj < 0.05 & log2FoldChange > 1.5)                            #ÕŹ│FLCAµ»öSLCAõĖŖĶ░āńÜäÕ¤║ÕøĀ
down_DEG <- subset(res_data, padj < 0.05 & log2FoldChange < -1.5)                         #ÕŹ│FLCAµ»öSLCAõĖŗĶ░āńÜäÕ¤║ÕøĀ
write.csv(res_data, paste("step5-",
                          paste(treatmentgroup, collapse="-"),
                          "-" ,
                          paste(controlgroup, collapse="-"), 
                          "-all.csv", 
                          sep=""))
write.csv(up_DEG, paste("step5-", 
                        paste(treatmentgroup, collapse="-"), 
                        "-" ,
                        paste(controlgroup, collapse="-"), 
                        "-up.csv", 
                        sep=""))
write.csv(down_DEG, paste("step5-", 
                          paste(treatmentgroup, collapse="-"),
                          "-" ,
                          paste(controlgroup, collapse="-"),
                          "-down.csv", 
                          sep=""))


res_data$change = ifelse(res_data$padj < 0.05 & abs(res_data$log2FoldChange) >1.5 , ifelse(res_data$log2FoldChange> 1.5 ,'Up','Down'),'Stable')
head(res_data)
#########################################################################################################
#########################                                              ##################################
#########################          STEP 6:fcshrink & MA plot           ##################################
#########################                                              ##################################
#########################################################################################################

#library(apeglm)
#resultsNames(dds_norm)
#res_shrink <- lfcShrink(dds_norm, coef= "condition_Leaf_vs_Flower", type="apeglm")                               #µ£ĆµÄ©ĶŹÉapeglmń«Śµ│Ģ;µĀ╣µŹ«resultsNames(dds)ńÜäń¼¼5õĖ¬ń╗┤Õ║”’╝īcoef=5’╝īõ╣¤ÕÅ»ńø┤µÄźcondition_LCE_vs_LCA"µīćÕ«Üń╗┤Õ║”
#pdf("../step6-MAplot.pdf", width = 6, height = 6)
#plotMA(res_shrink, ylim=c(-10,10), alpha=0.1, main="MA plot: ")
#dev.off()

#########################################################################################################
#########################                                              ##################################
#########################               STEP 7:ńü½Õ▒▒ÕøŠ                  ##################################
#########################                                              ##################################
#########################################################################################################

library(ggplot2)
pdf(paste("step7-", 
          paste(treatmentgroup, collapse="-"),
          "-" ,
          paste(controlgroup, collapse="-"),
          "-volcano.pdf", sep=""), 
    width = 6.13, height = 5.18)
p <- ggplot(
  # µĢ░µŹ«ŃĆüµśĀÕ░äŃĆüķó£Ķē▓
  res_data, aes(x =log2FoldChange , y = -log10(padj), colour=change)) +
  geom_point(alpha=0.6, size=0.6) +
  scale_color_manual(values=c("#546de5","black", "#ff4757","#d2dae2"))+
  # ĶŠģÕŖ®ń║┐
  geom_vline(xintercept=c(-1.5,1.5),lty=1,col="black",lwd=0.4) +
  geom_hline(yintercept =1.3,lty=1,col="black",lwd=0.4) +                            #ÕÅŹÕ»╣µĢ░,õ╗ŻĶĪ©0.05ńÜäń║┐,log10(0.05) =-1.3 
  # ÕØÉµĀćĶĮ┤
  labs(x="log2(fold change)",
       y="-log10 (padj)")+
  theme_bw()+
  # ÕøŠõŠŗ
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        legend.title = element_blank()) #õĖ╗µ¼ĪńĮæµĀ╝ń║┐ÕØćõĖ║ń®║ńÖĮ
print(p+xlim(-10, 10)+ylim(0, 10))
#p+xlim(-10, 10)+ylim(0, 10)
#p+scale_x_continuous(limits=c(5, 20))+scale_y_continuous(limits=c(0, 50))
#p + expand_limits(x = c(10, 10), y = c(-10, 30))
dev.off()
}}

```


##Co-Expression Analysis
###
```{r eval=FALSE}
library(pheatmap)
RNAexpr <- read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/CanA_fpkm.csv')
CYP_lst = read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/Sesquiterpenon_CYP.csv',header = T,sep=',')

idx = which(RNAexpr$Gene_ID %in% CYP_lst$ID)
RNAexpr <- RNAexpr[idx,]

RNAexpr$Flower1 <- rowMeans(RNAexpr[ , c(2, 3)], na.rm=TRUE)
RNAexpr$Flower2 <- rowMeans(RNAexpr[ , c(4, 5)], na.rm=TRUE)
RNAexpr$Flower3 <- rowMeans(RNAexpr[ , c(6, 7)], na.rm=TRUE)
#RNAexpr$Flower <- rowMeans(RNAexpr[ , 2:7], na.rm=TRUE)

RNAexpr$Leaf1 <- rowMeans(RNAexpr[ , c(8, 9)], na.rm=TRUE)
RNAexpr$Leaf2 <- rowMeans(RNAexpr[ , c(10, 11)], na.rm=TRUE)
RNAexpr$Leaf3 <- rowMeans(RNAexpr[ , c(13, 14)], na.rm=TRUE)
#RNAexpr$Leaf <- rowMeans(RNAexpr[ , 8:14], na.rm=TRUE)

RNAexpr$Root1 <- rowMeans(RNAexpr[ , c(15, 16)], na.rm=TRUE)
RNAexpr$Root2 <- RNAexpr[ , 17]
RNAexpr$Root3 <- rowMeans(RNAexpr[ , c(18, 19)], na.rm=TRUE)
#RNAexpr$Root <- rowMeans(RNAexpr[ , 15:19], na.rm=TRUE)

#RNAexpr$Stem <- rowMeans(RNAexpr[ , c(20, 21)], na.rm=TRUE)

RNAexpr = RNAexpr[,c(-2:-19)]
#RNAexpr = RNAexpr[,c(-2:-19)]

rownames(RNAexpr) <- RNAexpr$Gene_ID
expr_dat <- RNAexpr[,-1]

#bait_gene <- as.numeric(expr_dat['CaA11G00959', ])
bait_gene <- as.numeric(expr_dat['TPS16', ])

gene_cor <- apply(expr_dat,1,function(x) cor(x,bait_gene,method = 'pearson'))

top_genes_idx <- order(gene_cor, decreasing = TRUE)[1:25]

top_genes <- data.frame(
  Gene_ID = rownames(expr_dat)[top_genes_idx],
  Correlation = gene_cor[top_genes_idx]
)

print(top_genes)
#write_csv(top_genes,"/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/top_cpr_genes.csv")
top_expr_data <- expr_dat[rownames(expr_dat) %in% top_genes$Gene_ID, ]
top_expr_data <- top_expr_data[match(top_genes$Gene_ID, rownames(top_expr_data)), ]

z_score_expr <- t(scale(t(log2(top_expr_data + 1))))

pheatmap(z_score_expr, cluster_rows = FALSE, cluster_cols = FALSE, 
         color = colorRampPalette(c("blue", "white", "red"))(50))

#top_gene_expr <- expr_data[names(top_genes), ]

#################################### for CPR ###################
RNAexpr <- read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/CanA_fpkm.csv')
CPR_lst = read.csv("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/CPR.csv")
idx = which(RNAexpr$Gene_ID %in% CPR_lst$ID)
RNAexpr <- RNAexpr[idx,]

RNAexpr$Flower1 <- rowMeans(RNAexpr[ , c(2, 3)], na.rm=TRUE)
RNAexpr$Flower2 <- rowMeans(RNAexpr[ , c(4, 5)], na.rm=TRUE)
RNAexpr$Flower3 <- rowMeans(RNAexpr[ , c(6, 7)], na.rm=TRUE)

RNAexpr$Leaf1 <- rowMeans(RNAexpr[ , c(8, 9)], na.rm=TRUE)
RNAexpr$Leaf2 <- rowMeans(RNAexpr[ , c(10, 11)], na.rm=TRUE)
RNAexpr$Leaf3 <- rowMeans(RNAexpr[ , c(13, 14)], na.rm=TRUE)

RNAexpr$Root1 <- rowMeans(RNAexpr[ , c(15, 16)], na.rm=TRUE)
RNAexpr$Root2 <- RNAexpr[ , 17]
RNAexpr$Root3 <- rowMeans(RNAexpr[ , c(18, 19)], na.rm=TRUE)


RNAexpr = RNAexpr[,c(-2:-19)]

rownames(RNAexpr) <- RNAexpr$Gene_ID
expr_dat <- RNAexpr[,-1]

bait_gene <- as.numeric(expr_dat['CaA11G00959', ])

gene_cor <- apply(expr_dat,1,function(x) cor(x,bait_gene,method = 'pearson'))

top_genes_idx <- order(gene_cor, decreasing = TRUE)[1:25]

top_genes <- data.frame(
  Gene_ID = rownames(expr_dat)[top_genes_idx],
  Correlation = gene_cor[top_genes_idx]
)

print(top_genes)

top_expr_data <- expr_dat[rownames(expr_dat) %in% top_genes$Gene_ID, ]
top_expr_data <- top_expr_data[match(top_genes$Gene_ID, rownames(top_expr_data)), ]
z_score_expr <- t(scale(t(log2(top_expr_data + 1))))

pheatmap(z_score_expr, cluster_rows = FALSE, cluster_cols = FALSE, 
         color = colorRampPalette(c("blue", "white", "red"))(50))


#For TPS

RNAexpr <- read.csv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/CanA_fpkm.csv')
TPS_lst = read.csv("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/TPS_ID.csv")
idx = which(RNAexpr$Gene_ID %in% TPS_lst$ID)
RNAexpr <- RNAexpr[idx,]

RNAexpr$Flower1 <- rowMeans(RNAexpr[ , c(2, 3)], na.rm=TRUE)
RNAexpr$Flower2 <- rowMeans(RNAexpr[ , c(4, 5)], na.rm=TRUE)
RNAexpr$Flower3 <- rowMeans(RNAexpr[ , c(6, 7)], na.rm=TRUE)

RNAexpr$Leaf1 <- rowMeans(RNAexpr[ , c(8, 9)], na.rm=TRUE)
RNAexpr$Leaf2 <- rowMeans(RNAexpr[ , c(10, 11)], na.rm=TRUE)
RNAexpr$Leaf3 <- rowMeans(RNAexpr[ , c(13, 14)], na.rm=TRUE)

RNAexpr$Root1 <- rowMeans(RNAexpr[ , c(15, 16)], na.rm=TRUE)
RNAexpr$Root2 <- RNAexpr[ , 17]
RNAexpr$Root3 <- rowMeans(RNAexpr[ , c(18, 19)], na.rm=TRUE)


RNAexpr = RNAexpr[,c(-2:-19)]

rownames(RNAexpr) <- RNAexpr$Gene_ID
expr_dat <- RNAexpr[,-1]


write.csv(expr_dat,"/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/1-CoExpr/TPS_expr.csv",row.names = T)
library(dplyr)
library(tidyr)


# 1. Ķ»╗ÕÅ¢µĢ░µŹ« (õĮ┐ńö© check.names = FALSE õ┐ØµīüÕÄ¤Õ¦ŗÕłŚÕÉŹ’╝īÕ”é Stem_1)
expr_dat <- read.csv("TPS_expr.csv", row.names = 1, check.names = FALSE)
library(dplyr)
library(tidyr)
#

```


##WGCNA

Weighted Gene Co-Expression Network Analysis (WGCNA) has been widely used to create co-expression networks, grouping genes with similar expression pattern in clusters and relating these cluster with phenotypic characterics.
```{r eval=FALSE}
# BiocManager::install("WGCNA")
# BiocManager::install("GO.db")
# BiocManager::install("impute")
# BiocManager::install("preprocessCore")
library(WGCNA)
library(reshape2)
options(stringsAsFactors = FALSE) #Setting string not as factor
enableWGCNAThreads()                                         #Enable multiple process

RNAexpr <- read.csv("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/CanA_fpkm.csv",sep=",",header = T) #gene expression datafile

RNAexpr$Flower1 <- rowMeans(RNAexpr[ , c(2, 3)], na.rm=TRUE)
RNAexpr$Flower2 <- rowMeans(RNAexpr[ , c(4, 5)], na.rm=TRUE)
RNAexpr$Flower3 <- rowMeans(RNAexpr[ , c(6, 7)], na.rm=TRUE)

RNAexpr$Leaf1 <- rowMeans(RNAexpr[ , c(8, 9)], na.rm=TRUE)
RNAexpr$Leaf2 <- rowMeans(RNAexpr[ , c(10, 11)], na.rm=TRUE)
RNAexpr$Leaf3 <- rowMeans(RNAexpr[ , c(13, 14)], na.rm=TRUE)

RNAexpr$Root1 <- rowMeans(RNAexpr[ , c(15, 16)], na.rm=TRUE)
RNAexpr$Root2 <- RNAexpr[ , 17]
RNAexpr$Root3 <- rowMeans(RNAexpr[ , c(18, 19)], na.rm=TRUE)

RNAexpr = RNAexpr[,c(-2:-19)]
RNAexpr = RNAexpr[,-7]

rownames(RNAexpr)<-RNAexpr[,1]

#fpkm = RNAexpr[,-1]

#countData<-RNAexpr[,-1]

DEGsexpr <- RNAexpr[RNAexpr$Gene_ID %in% unique_degs, ]
rownames(DEGsexpr)=DEGsexpr[,1]

fpkm = DEGsexpr[,-1]

samplename <- c(rep("Stem",2),rep("Flower",3),rep("Leaf",2),rep("Root",3))
samplename

datTraits = data.frame(gsm=names(fpkm),
                       sampletype=samplename 
                       )                                     
datTraits

rownames(datTraits)=datTraits[,1]

RNAseq_voom <- fpkm

dim(RNAseq_voom)
WGCNA_matrix = t(RNAseq_voom[order(apply(RNAseq_voom,1,mad), decreasing = T)[1:12576],])          

dim(WGCNA_matrix)
#Data Check before proceeding; if any problematic samples/genes, the matrix is filtered
datExpr <- WGCNA_matrix                                      # top 29364 mad genes
datExpr[1:10,1:4] #Sample the first 11 rows (the number of samples)and the first 4 columns to inspect to check

gsg = goodSamplesGenes(datExpr)                             #gsg data check/filteration

if (!gsg$allOK) {
        datExpr = datExpr[gsg$goodSamples, gsg$goodGenes]
}
dim(datExpr)

#####################STEP 2: Check outliar==========================================================
datExpr_tree<-hclust(dist(datExpr), method = "average") #Group data in a dendogram to check outliers
par(mar = c(0,5,2,0))

plot(datExpr_tree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 2,
     cex.axis = 1, cex.main = 1,cex.lab=1)

#dim(datExpr)
nGenes = ncol(datExpr)  #the number of genes
nSamples = nrow(datExpr) #the number of samples

sample_colors <- numbers2colors(as.numeric(factor(datTraits$sampletype)),
                                colors = c("cyan",
                                           "red",
                                           "green3",
                                           "gray"),
                                signed = FALSE)                                                  
                                   
par(mar = c(1,4,3,1),cex=0.8)

plotDendroAndColors(datExpr_tree, sample_colors,
                    groupLabels = colnames(sample),
                    cex.dendroLabels = 0.8,
                    marAll = c(1, 4, 3, 1),
                    cex.rowText = 0.01,
                    main = "Sample dendrogram and sampletype heatmap")         

#ensures that your trait data is correctly aligned with your expression data
sampleNames = rownames(datExpr)
traitRows = match(sampleNames, datTraits$gsm)
rownames(datTraits) = datTraits[traitRows,1] #=datTraits[,1,traitRows]=datTraits[1:16, 1]
# if (sum(is.na(traitRows)) == 0) {
#     # Update row names in datTraits to match those in datExpr
#     rownames(datTraits) = datTraits[traitRows, 1]
# } else {
#     warning("Some sample names could not be matched.")
# }

################STEP 3: Identify optimal soft-thresholding power##################################
powers = c(c(1:10), seq(from = 20, to=30, by=2))                
sft = pickSoftThreshold(datExpr,
                        powerVector = powers,
                        networkType = "signed", 
                        RsquaredCut = 0.7,
                        moreNetworkConcepts = TRUE,
                        blockSize = 2000,
                        verbose = 5)
#sft = pickSoftThreshold(datExpr, RsquaredCut = 0.8,powerVector = powers, verbose = 5) #r>0.9 r>0.85
sft
png("step2-beta-value-Fina1.png",width = 800,height = 600)       # Plot the results
par(mfrow = c(1,2));
cex1 = 0.9;

#Index the scale free topology adjust as a function of the power soft thresholding.
p <- plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));                 

text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
abline(h=0.70,col="green")      #This line corresponds to use a cut-off R┬▓ of h

#Connectivity mean as a function of soft power thresholding
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))                          
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
dev.off()

best_beta=sft$powerEstimate
#blockwiseModules() for construction the TOM (topological overlap matix)
net = blockwiseModules(
    datExpr,
    power = sft$powerEstimate,
    networkType = "signed",
    TOMType = "signed",
    #detectCutHight
    minModuleSize = 30,
    maxBlockSize = 6000,
    #Module Adjustment
    reassignThreshold = 0,
    mergeCutHeight = 0.25,
    numericLabels = TRUE,
    pamRespectsDendro = FALSE,
    #Generate TOM file
    saveTOMs = TRUE,
    saveTOMFileBase = "Fina1-FPKM-TOM",
    verbose = 3
    )

#Convert labels to colors for plotting
table(net$colors)
mergedColors = labels2colors(net$colors)                                
table(mergedColors)
write.table(table(mergedColors),
            quote = F,
            row.names = T,
            append = F,
file = "/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/step4-mod_Gene_Number-Fina1.txt",
      col.names = F)
#Plot the dendrogram and the module colors undreneath
p <- plotDendroAndColors(net$dendrograms[[1]], 
                         mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE,
                    hang = 0.03,
                    addGuide = TRUE,
                    guideHang = 0.05)


#Relate Module (cluster) Assignments to Treatment Groups.
design1=model.matrix(~0+datTraits$sampletype)  

colnames(design1)= levels(as.factor(datTraits$sampletype))             

moduleColors <- labels2colors(net$colors)
#Get Module Eigengenes (hypothetical central gene) per cluster         
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes 
#Record modules so similar modules are next to each other
MEs = orderMEs(MEs0)                                               

ModuleName=unique(substring(colnames(MEs),3))     

#correlation between Module and Traint/Phenotypes
moduleTraitCor = cor(MEs, design1 , use = "p")

moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
sizeGrWindow(10,6)       
#Displaying correlations and its p-value
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "")

dim(textMatrix) = dim(moduleTraitCor)

pdf("step5-Module-sample-relationships-Fina1.pdf", width = 10, height = 15)          

par(mar = c(6, 10, 1, 1))
#Displaying the correlation values in a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels =  colnames(design1),
               yLabels = names(MEs),
               ySymbols = ModuleName,
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
dev.off()

png("step5-Module-adjacency-heatmap-Fina1.png",height = 800,width = 600)          
plotEigengeneNetworks(MEs, "Eigengene adjacency heatmap",plotDendrograms = F,
                      marDendro = c(4,4,2,4))
dev.off()

setwd("/home/zhuandong/03-Cangulatus/01-Cangulatus_A/03-Transcriptome_analysis/wgcna/Root_green_data")#*

SAMP = as.data.frame(design1[,1])                                           

names(SAMP) = "Root"#*
module = "brown" #*
which.module="brown";#*

moduleColors <- labels2colors(net$colors)
MEs = moduleEigengenes(datExpr, moduleColors)$eigengenes                        

MET = orderMEs(cbind(MEs, SAMP))                                                   

sizeGrWindow(5,7.5);
par(cex = 0.9)
pdf(file =paste("step7-",
                paste(names(SAMP), collapse="-"),
                "-Eigengene-adjacency-and-dendrogram-Fina1.pdf",
                sep=""),
    width = 10, height = 12)                                                       
plotEigengeneNetworks(MET, "",
                      marDendro = c(0,4,1,2),
                      marHeatmap = c(3,4,1,2),
                      cex.lab = 0.7,
                      xColorWidth = 1 * strheight("M"),
                      xLabelsAngle
                      = 90)
dev.off()
sizeGrWindow(6,6);
par(cex = 1.0)
png(filename =  paste("step7-",
                paste(names(SAMP), collapse="-"),
                "-Eigengene-dendrogram-hclust-Fina1.png",
                sep=""),
    width = 800,height = 600)                                                     
plotEigengeneNetworks(MET, "Eigengene dendrogram",
                      marDendro = c(0,4,2,0),
                      plotHeatmaps = FALSE)
dev.off()

par(cex = 1.0)
pdf(file =paste("step7-",
                paste(names(SAMP), collapse="-"),
                "-Eigengene-adjacency-heatmap-Fina1.pdf",
                sep=""),
    width = 10, height = 12)                                                     #┬╗├”├¢├å├æ├╣┬▒┬Š┬╗├▓├É├ö├Ś┬┤┬Ī┬ó├å┬Ę├Å┬Ą├ō├½├ä┬Ż┬┐├®┬Ą├ä├ü├Ü┬Į├ō┬Š├ś├Ģ├│├ł├ł├Ź┬╝┬Ż┬©├Ś┬ó├Æ├ó┬Ż┬║┬┤├ŗ├Ź┬╝┬Į┬½┬Ė┬▓┬Ė├ć├Ŗ├Ę├Ś┬┤├Ź┬╝┬Ż┬®
plotEigengeneNetworks(MET, "Eigengene adjacency heatmap",
                      marHeatmap = c(3,4,2,2),
                      plotDendrograms = FALSE,
                      #cex.lab = 0.7,
                      cex.lab.x = 0.7,
                      cex.lab.y = 0.8,
                      xColorWidth = 1 * strheight("M"),
                      xLabelsAngle = 90)
dev.off()
file.remove('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/step8-All_Gene_KME-Fina1.txt')
for(modules in substring(colnames(MEs),3)){
    if(modules == "grey") next
    ME=as.data.frame(MEs[,paste("ME",modules,sep="")])
    colnames(ME)=modules
    datModExpr=datExpr[,moduleColors==modules]
    datKME = signedKME(datModExpr, ME)
    datKME=cbind(datKME,rep(modules,length(datKME)))
    write.table(datKME,quote = F,row.names = T,append = T,file = "/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/step8-All_Gene_KME-Fina1.txt",col.names = T)
}

modNames = substring(names(MEs), 3)                                                # ├ä┬Ż┬┐├®┬Ą├ä├ā├╗┬│├å┬Ż┬©├æ├Ģ├ē┬½┬Ż┬®
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"))         
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))
names(geneModuleMembership) = paste("MM", modNames, sep="")
names(MMPvalue) = paste("p.MM", modNames, sep="")
geneModuleMembership[1:4,1:4]
geneTraitSignificance = as.data.frame(cor(datExpr, SAMP, use = "p"))               #

GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))
names(geneTraitSignificance) = paste("GS.", names(SAMP), sep="")
names(GSPvalue) = paste("p.GS.", names(SAMP), sep="")
column = match(module, modNames)
moduleGenes = moduleColors==module

sizeGrWindow(7, 7)
png(filename =  paste("step8-",
                      paste(module, collapse="-"),
                      "-membership-gene_significance-Fina1.png",
                      sep=""),
    width = 800,height = 600)
par(mfrow = c(1,1))
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = paste("Gene significance for ", names(SAMP),sep=""), #*
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col =module)
dev.off()
moduleColors <- labels2colors(net$colors)

#connet=abs(cor(datExpr,use="p"))^sft$powerEstimate  #├Ä├×├Å├▓├Ź├Ė├é├¦unsigned
#connet=(1+cor(datExpr,use="p")/2) ^sft$powerEstimate #├ō├É├Å├▓├Ź├Ė├é├¦sigined
#Alldegrees1=intramodularConnectivity(connet, moduleColors)
Alldegrees1=intramodularConnectivity.fromExpr(datExpr,
                                              moduleColors,
                                              networkType = "signed",
                                              power =sft$powerEstimate,
                                              ignoreColors = FALSE)
rownames(Alldegrees1)=colnames(datExpr)
head(Alldegrees1)
#file.remove('../step8-All_moduleGene_Connectivity-Fina1.txt')
write.table(Alldegrees1,quote = F,row.names = T,append = F,file = "/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/step8-All_moduleGene_Connectivity-Fina1.txt",col.names = T)
GS1 = as.numeric(cor(SAMP,datExpr, use="p"))                            #├ä┬│┬Ė├Č├æ├╣┬▒┬Š┬║├Ź├ŗ├╣├ō├É┬╗├╣├Æ├▓┬Ą├ä├Å├Ā┬╣├ś├É├ö┬Ż┬¼├Å├Ā┬╣├ś├É├ö┬Ą├ä├Ģ├Į┬Ė┬║┬┤├│├É┬Ī┬▒├Ł├Ŗ┬Š┬Ė├ā├æ├╣┬▒┬Š├¢├É┬╗├╣├Æ├▓┬▒├Ł┬┤├»├ü┬┐┬Ą├ä┬Ė├¤┬Ą├Ź
GeneSignificance=abs(GS1)                                              #├¢┬╗├ō├ā┬║┬»├Ŗ├Į┬╝├å├ŗ├Ż├Å├Ā┬╣├ś├É├ör├¢┬Ą┬Ż┬¼├ā┬╗├ō├É├ō├ā┬║┬»├Ŗ├Į┬Į├Ė├É├ÉT┬╝├¼├æ├®┬╝├å├ŗ├ŻP├¢┬Ą┬Ż┬¼┬╗├╣├Æ├▓├Å├ö├¢├Ė├É├ö├¢┬Ą├Ä┬¬├Å├Ā┬╣├ś├É├ö├¢┬Ą┬Ą├ä┬Š├Ė┬Č├ö├¢┬Ą┬Ż┬¼P├¢┬Ą=┬╗├╣├Æ├▓├Å├ö├¢├Ė├É├ö├¢┬Ą┬Ą├ä┬Č├ö├Ŗ├Į┬Ī┬Ż
colorlevels=unique(moduleColors)
sizeGrWindow(9,6)
png(filename =  paste("step8-",
                      paste(module, collapse="-"),
                      "-Gene_Significance-Connectivity-Fina1.png",
                      sep=""),
    width = 800,height = 600)
#pdf("step8-Gene_Significance-Connectivity-Fina1.pdf", width = 10, height = 15)
par(mfrow=c(2,as.integer(0.5+length(colorlevels)/2)))
#par(mar = c(4,5,3,1))
par(mar = c(1,0.1,1,1))
for (i in c(1:length(colorlevels)))
{
    whichmodule=colorlevels[[i]];
    restrict1 = (moduleColors==whichmodule);
    verboseScatterplot(Alldegrees1$kWithin[restrict1],
                       GeneSignificance[restrict1], col=moduleColors[restrict1],
                       main=whichmodule,
                       xlab = "Connectivity", ylab = "Gene Significance", abline = TRUE)
}
dev.off()
datKME=signedKME(datExpr, MEs, outputColumnName="MM.")
head(datKME)                                                                #├Å├ö├Ŗ┬Š├Ŗ├Į┬Š├Ø┬┐├▓┬Ą├ä├ć┬░┬╝┬Ė├É├É
FilterGenes1= abs(GS1) > .8 & abs(datKME[,which(names(datKME)== paste("MM.",module,sep=""))])>.9 #*                    #┬Ė├É├É├ŗ├ł┬ż├æ├╣┬▒┬Š├¢├É┬▒├Ł┬┤├»├ü┬┐┬Į├Å┬Ė├¤┬Ą├ä┬Ė├É├É├ŗ├ł┬ż├ä┬Ż┬┐├®┬Ą├äHub┬╗├╣├Æ├▓        #abs(GS1)> .9 ┬┐├ē├Æ├ö┬Ė├╣┬Š├Ø├Ŗ┬Ą┬╝├Ŗ├ć├®┬┐├Č┬Ą├Ę├Ģ├╗┬▓├Ä├Ŗ├Į(┬╣├Ö├Ź├Ė├Ŗ├ć>.2┬Ż┬¼├Æ┬╗┬░├Ż├ē├©├¢├ā>.8)┬Ż┬¼├Æ├▓├Ä┬¬├Ŗ├ć├ä┬│┬Ė├Č├æ├╣┬▒┬Š┬Ą├ä┬╗├╣├Æ├▓┬Ė├¤├ō├Ü├ŗ├╣├ō├É├æ├╣┬▒┬Š┬Ė├ā┬╗├╣├Æ├▓┬Ą├ä├å┬Į┬Š├╣┬▒├Ł┬┤├»├ü┬┐(>0 ┬Š├Ź├ŗ┬Ą├ā├Ę┬Ė├¤├ō├Ü├å┬Į┬Š├╣┬▒├Ł┬┤├»├ü┬┐)┬Ż┬¼┬Ą┬½├Ŗ├ć├ö├Ü├ä┬│┬Ė├Č├æ├╣┬▒┬Š├¢├É┬Ą├ä┬▒├Ł┬┤├»├ü┬┐┬Į├Å┬Ė├¤┬▓┬╗├Æ┬╗┬Č┬©├Ŗ├ćhub genes┬Ż┬¼├¢├Ę├Æ┬¬┬┐┬┤KEM├¢┬Ą┬Ī┬Żabs(datKME$MM.modname)>.8 ├¢├ü├ē├Ö┬┤├│├ō├Ü >0.8 HubGene: |kME| >=├Ż├É├¢┬Ą┬Ż┬©0.8┬Ż┬®
table(FilterGenes1)
FilterGenes2=data.frame(colnames(datExpr),FilterGenes1)
library(dplyr)
FilterGenes3=filter(FilterGenes2,FilterGenes1=="TRUE")
write.table(FilterGenes3,quote = F,row.names = T,append = F,
            file = paste("step8-",
                         paste(module, collapse="-"),
                         "-KME-hub_Gene-Fina1.txt",
                         sep=""),
            col.names = F)
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
geneTree = net$dendrograms[[1]];
dissTOM = 1-TOMsimilarityFromExpr(datExpr,
                                  networkType = "signed",
                                  power = sft$powerEstimate,
                                  TOMType="signed");
nSelect = 400
set.seed(10);                                                           # For reproducibility, we set the random seed
select = sample(nGenes, size = nSelect);
selectTOM = dissTOM[select, select];
selectTree = hclust(as.dist(selectTOM), method = "average")             # ├ā┬╗├ō├É┬╝├▓┬Ą┬ź┬Ą├ä┬Ę┬Į┬Ę┬©┬┐├ē├Æ├ö┬Į┬½┬Š├ø├Ć├Ā├Ŗ├Ę├Å├×├¢├å├Ä┬¬┬╗├╣├Æ├▓├Ś├ō┬╝┬»┬Ż┬¼├Æ├▓┬┤├ŗ├Ä├Æ├ā├ć┬▒├ś├É├½┬Č├ö├ŗ├”┬╗├║┬Ą├ä┬╗├╣├Æ├▓├Ś├ō┬╝┬»├¢├ś├É├é┬Š├ø├Ć├Ā┬Ī┬Ż
selectColors = moduleColors[select];
sizeGrWindow(9,9)                                                       # Open a graphical window
plotDiss = selectTOM^7;                                                 #├Ź┬©┬╣├Į├Ś┬¬┬╗┬╗┬║├Ź├ō├É├É┬¦┬Ą├ś┬Ė├╝┬Ė├ädissTOM ┬Ą├ä├ā├Ø┬Ż┬¼├Ć├Į├ł├¦10┬Ż┬¼├Ŗ┬╣┬╗├”├Ź┬╝├É├ģ├Å┬ó┬Ė├╝┬╝├ō┬Ę├Ī┬Ė┬╗┬Ż┬¼├Ŗ┬╣├¢├É┬Ą├ł├ć┬┐┬Č├ł┬Ą├ä├ü┬¼┬Į├ō├ö├Ü├ł├ł├Ź┬╝├¢├É┬Ė├╝├ā├Ę├Å├ö┬Ī┬Ż
diag(plotDiss) = NA;                                                    # the color palette; setting the diagonal to NA also improves the clarity of the plot
png("../step9-Select-Network-heatmap-Fina1.png",
    width = 800,height = 600)
TOMplot(plotDiss,
        selectTree,
        selectColors,
        main = "Network heatmap plot, selected genes")
dev.off()
probes = colnames(datExpr)                                              # probe┬Š├Ź├Ŗ├ć┬╗├╣├Æ├▓├ā├╗
inModule = (moduleColors==module);
modProbes = probes[inModule];
write.table(modProbes,quote = F,
            row.names = T,
            append = F,
            file =paste("step9-",
                        paste(module, collapse="-"),
                        "-module_Gene_Name-Fina1.txt",
                        sep=""),
            col.names = F)
TOM = TOMsimilarityFromExpr(datExpr,
                            networkType = "signed",
                            power = sft$powerEstimate,
                            TOMType="signed");                         

#module = "lightgreen";#*                                               # Select module
probes = colnames(datExpr)                                             # probe┬Š├Ź├Ŗ├ć┬╗├╣├Æ├▓├ā├╗
inModule = (moduleColors==module);
modProbes = probes[inModule];                                          #├ī├Ī├ł┬Ī├¢┬Ė┬Č┬©├ä┬Ż┬┐├®┬Ą├ä┬╗├╣├Æ├▓├ā├╗
modTOM = TOM[inModule, inModule];                                      #├æ┬Ī├ö├▒├ä┬Ż┬┐├®├ä├Ü┬╗├╣├Æ├▓├Å├Ā├ō┬”┬Ą├ä├Ź├ś├å├ŗ├¢├ś┬Ą├ŠTOM├¢┬Ą
dimnames(modTOM) = list(modProbes, modProbes)
IMConn <- softConnectivity(datExpr[, modProbes])
head(IMConn)
all=data.frame(modProbes,IMConn)
order=all[order(all[,2],decreasing = T),]
write.table(order,quote = F,row.names = F,append = F,file = paste("step9-",
                                         module,
                                         "-paiming-Fina1.txt",
                                         sep=""),col.names = F)

vis = exportNetworkToVisANT(modTOM,
                            file = paste("step9-",
                                         module,
                                         "-VisANTInput-Fina1.txt",
                                         sep=""),
                            weighted = TRUE,
                            threshold = 0)                             

cyt = exportNetworkToCytoscape(modTOM,
                               edgeFile = paste("step9-",
                                                paste(module, collapse="-"),
                                                "-CytoscapeInput-edges-Fina1.txt",
                                                sep=""),
                               nodeFile = paste("step9-",
                                                paste(module, collapse="-"),
                                                "-CytoscapeInput-nodes-Fina1.txt",
                                                sep=""),
    weighted = TRUE,
    threshold = 0.25,
    nodeNames = modProbes,
    nodeAttr = moduleColors[inModule]
);

cyt_data = read.table('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/06-Transcript_Analysis/03-WGCNA/step9-brown-CytoscapeInput-edges-Fina1.txt',
                      header = T,
                      comment.char = "", 
                      check.names=F)
```


##Pathway
```{r}
rm(list = ls())
setwd('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/05-Transcript_Analysis/Pathway')
#install.packages("gplots")
#install.packages("pheatmap")
library(gplots)
library(pheatmap)

data <- read.csv("pathway_genes_expr.txt",header = T,sep='\t',quote = '', fill = TRUE)
data
#a = c(#"all-tonglu",
      #"AACT","CMK","DXPR","DXPS","FPPS","GPPS","HDR","HDS","HMGR","HMGS","IDI","MCT","MDS","MVD","MVK","PMVK"
#      "STPS","B"
#      )
#for (NAME in a) {

#data<-read.csv(paste(paste
#                 (NAME, collapse="."), 
#                     ".csv", 
#                     sep=""),
#               head=T)
head(data)

row.names(data)<-data[,1]
row.names(data)
data<-data[,-1]
df1 <-as.matrix(data)
df1
#dat <- t(df1)
df2=log2(df1+1)#log??????ū╝???????ŪĘ?╚Ī?????╔Ė????į╝??????¾Ż¼ĘŪ▒??ļĪŻ
#df = df1[apply(df1, 1, function(x) sd(x)!=0),]#╚ź??NA??
NAME="LH3-3"
p = pheatmap(data, 
             #df2,
         show_rownames = T,
         show_colnames = T,
         border="white",# ?▀┐???╔½
         cellwidth = 80,#  ?▀┐?????
         cellheight=50,
         cluster_cols = F,
         cluster_rows= F, # ?ąŠ???
         #filename='cyan.fpkm6.png',#?????─╝???????
         filename=paste(paste(NAME, collapse="."), 
               ".pdf", 
               sep=""),
         fontsize_row =10, #???????─┤?ąĪ
         fontsize_col = 10,
         #height=10,  #????═╝Ų¼?─Ė▀Č?
         scale = "row",#?????ąĮ???z-score??ū╝????????Ū░??log??????ū╝???╦┐???scale="none",
         angle_col=0, #?▐Ė─║???????????ą▒??
         legend= T, # ╚ź??═╝??
         color = colorRampPalette(c("steelblue3", "white", "firebrick3"))(100),#??═╝??????╔½
         clustering_distance_rows = 'euclidean', 
         #clustering_method = 'complete',
)#}
p

```



```{r}
rm(list = ls())
setwd("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/04-Evolution_Analysis/01-GeneFam")

#plot genefam on Chrs.
#par(mfrow=c(2,2),mar=c(4,4,1,1)
plot(62:100, type = 'n',xlim = c(0,70),ylim = c(0,17500000),las=1, axes=F,xlab = '',ylab = '') 
#'type': what type of plot should be drawn, 'type='"n"' for no plotting, only axis;axes=F for personalized axis
axis(2,seq(0,18000000,by=2000000),las=1,col = 'black')  #axis function for personalized axis;side=2,setting y axis
#chr1
lines(x=c(0,0),y=c(0,17401811),lwd=2.5,col='black',type='h')  #type=h for histogram
#chr2
lines(x=c(3,3),y=c(0,16452266),lwd=2.5)
#chr3
lines(x=c(6,6),y=c(0,15164551),lwd=2.5)
#chr4
lines(x=c(9,9),y=c(0,15202774),lwd=2.5)
#chr5
lines(x=c(12,12),y=c(0,14778003),lwd=2.5)
#chr6
lines(x=c(15,15),y=c(0,14664815),lwd=2.5)
#chr7
lines(x=c(18,18),y=c(0,14769230),lwd=2.5)
#chr8
lines(x=c(21,21),y=c(0,14645955),lwd=2.5)
#chr9
lines(x=c(24,24),y=c(0,14326162),lwd=2.5)
#chr10
lines(x=c(27,27),y=c(0,14315424),lwd=2.5)
#chr11
lines(x=c(30,30),y=c(0,14033347),lwd=2.5)
#chr12
lines(x=c(33,33),y=c(0,14208336),lwd=2.5)
#chr13
lines(x=c(36,36),y=c(0,13967967),lwd=2.5)
#chr14
lines(x=c(39,39),y=c(0,13931248),lwd=2.5)
#chr15
lines(x=c(42,42),y=c(0,13998627),lwd=2.5)
#chr16
lines(x=c(45,45),y=c(0,13564716),lwd=2.5)
#chr17
lines(x=c(48,48),y=c(0,12861505),lwd=2.5)
#chr18
lines(x=c(51,51),y=c(0,12734079),lwd=2.5)
#chr19
lines(x=c(54,54),y=c(0,12381310),lwd=2.5)
#chr20
lines(x=c(57,57),y=c(0,12505263),lwd=2.5)
#chr21
lines(x=c(60,60),y=c(0,12244057),lwd=2.5)
#chr22
lines(x=c(63,63),y=c(0,11992425),lwd=2.5)
#chr23
lines(x=c(66,66),y=c(0,11737041),lwd=2.5)

gene_onbar=read.table('CanA_CYP_TPS_chr_v2.txt',header = 1)
chr_pos = c()
#set the horizontal distance between the genes and its chromosome to 0.5
for (i in c(1:length(gene_onbar$chr))) {
  chr_pos[i]=gene_onbar$chr[i]*3-3
}

gene_onbar$chr_pos = chr_pos #add a colummn called chr_pos in gene_onbar

gene_onbar_red = subset(gene_onbar,color=='red')
gene_onbar_blue = subset(gene_onbar,color=='blue')
#gene_onbar_black = subset(gene_onbar,color=='black')
#gene_onbar_green = subset(gene_onbar,color=='green')

#subset function return subsets of vector, matrics or data frames which meet conditions

for (i in c(1:length(gene_onbar_blue$chr))) {
  lines(x=c(gene_onbar_blue$chr_pos[i],gene_onbar_blue$chr_pos[i]+2),y=c(gene_onbar_blue$gene_pos[i],gene_onbar_blue$gene_pos[i]),
        col=as.character(gene_onbar_blue$color[i])) #as.character convert a numeric object into a character subject.
}

for (i in c(1:length(gene_onbar_red$chr))) {
  lines(x=c(gene_onbar_red$chr_pos[i],gene_onbar_red$chr_pos[i]+2), y=c(gene_onbar_red$gene_pos[i],gene_onbar_red$gene_pos[i]),
        col=as.character(gene_onbar_red$color[i]))
}

#for (i in c(1:length(gene_onbar_black$chr))) {
#  lines(x=c(gene_onbar_black$chr_pos[i],gene_onbar_black$chr_pos[i]+1),y=c(gene_onbar_black$chr_pos[i],gene_onbar_black$chr_pos[i]),
#        col=as.character(gene_onbar_black$chr_pos[i]))
#}
#for (i in c(1:length(gene_onbar_green$chr))) {
#  lines(x=c(gene_onbar_green$chr_pos[i],gene_onbar_green$chr_pos[i]+1),y=c(gene_onbar_green$chr_pos[i],gene_onbar_green$chr_pos[i]),
#        col=as.character(gene_onbar_green$chr_pos[i]))
#}
#for (i in c(1:length(gene_onbar_orange$chr))) {
#  lines(x=c(gene_onbar_orange$chr_pos[i],gene_onbar_orange$chr_pos[i]+1),y=c(gene_onbar_orange$chr_pos[i],gene_onbar_orange$chr_pos[i]),
#        col=as.character(gene_onbar_orange$chr_pos[i]))
#}


```

##Haplotype Visualization
### Identity
```{r}
identity_df <- read.delim(
  file = '/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/window_100k_identities.tsv',
  header = T,
  sep = '\t'
  )

p <- ggplot(identity_df,aes(x=Identity)) +
  geom_histogram(binwidth=0.5,
                 fill='#2E86AB', color='black',
                 alpha = 0.8) +
  theme_minimal() +
  labs(title='Haplotype Identity Distribution',
       x='Identity (%)',
       y='No.blocks') +
  theme_classic(base_size = 6)+
  theme(
     axis.line = element_blank(),
     panel.background = element_blank(),
     panel.border = element_rect(colour = "black",fill = NA,linewidth = 0.5),
     axis.text.x = element_text(),
     panel.grid = element_blank(),
     plot.title = element_text(hjust = 0.5, vjust = 1, face = "bold")
   ) +
  scale_x_continuous(
    breaks = seq(65, 100, by = 2.5),
    limits = c(65, 100)) +
  scale_y_continuous(limits = c(0,800),expand = c(0,0),
                     breaks = seq(0,800,by=200))
  
p
```


###CaTPS16 on Chr20
```{r}
library(Gviz)
library(GenomicRanges)
options(ucscChromosomeNames=FALSE)
# ====================================
# 1. Õ«Üõ╣ēń╗śÕøŠÕÅéµĢ░ÕÆī Seqinfo
# ====================================
target_chrom <- "chr20"
start_coord <- 11323432
end_coord <- 11330260
chrom_length <- 12505263
custom_genome_name <- "Cangulatus" 

# ÕłØÕ¦ŗÕī¢ń®║ńÜäĶĮ©Ķ┐╣ÕłŚĶĪ©
all_tracks <- list()

# ÕłøÕ╗║ Seqinfo Õ»╣Ķ▒Ī (ÕīģÕÉ½µ¤ōĶē▓õĮōķĢ┐Õ║”õ┐Īµü»)
chrom_seqinfo <- Seqinfo(
  seqnames = target_chrom,
  seqlengths = chrom_length,
  genome = custom_genome_name 
)

# ====================================
# 2. Create Gviz Tracks
# ====================================

# --- Track #1: IdeogramTrack (Õģ©ķĢ┐µ¤ōĶē▓õĮōÕøŠ) ---
full_chrom_band <- data.frame(
  chrom = target_chrom,            
  chromStart = 1,        
  chromEnd = chrom_length,            
  name = "chr20_full",
  gieStain = "gneg" 
)
ideo_track <- IdeogramTrack(
  genome = custom_genome_name, 
  chromosome = target_chrom,
  name = " ",
  bands = full_chrom_band,
  seqinfo = chrom_seqinfo, 
  outline = TRUE,
  fontcolor = "black",
  cex = 0.8,
  showTitle = FALSE,
  showId = FALSE
)
all_tracks <- append(all_tracks, ideo_track) 

# --- Track #2: GenomeAxisTrack (ÕØÉµĀćĶĮ┤) ---
axis_track <- GenomeAxisTrack(
  name = "Genomic Position (bp)",
  add53 = TRUE, 
  col = "black",
  cex = 0.8
)
all_tracks <- append(all_tracks, axis_track) 

# --- Track #3: GeneRegionTrack (Õ¤║ÕøĀń╗ōµ×ä - Hap1) ---
# µ©Īµŗ¤µĢ░µŹ«’╝ÜĶ»Ęµø┐µŹóõĖ║µé©ńÜäÕ«×ķÖģÕ¤║ÕøĀµ│©ķćŖµĢ░µŹ«’╝ü
gene_annot_df <- data.frame(
  chromosome = target_chrom,
  start = c(11325432, 11326246, 11326570, 11326844,
            11327149, 11327633, 11328032), 
  end = c(11325957, 11326494, 11326708, 11327062,
          11327521,11327906,11328260),   
  strand = "-", 
  feature = c("exon", "exon", "exon","exon",
              "exon","exon","exon"),
  gene = "CaTPS16",
  transcript = "CaTPS16.T1",
  Haplotype = "Hap1" 
)

gene_track_hap1 <- GeneRegionTrack(
  gene_annot_df,
  chromosome = target_chrom,
  genome = custom_genome_name,
  seqinfo = chrom_seqinfo, 
  name = " ",
  fill = "lightblue",
  col = NULL,
  stacking = "squish", 
  featureAnnotation = "gene"
)
all_tracks <- append(all_tracks, gene_track_hap1) 

# --- Track #4: AnnotationTrack (ÕÅśÕ╝éõĮŹńé╣) ---
# µ©Īµŗ¤µĢ░µŹ«’╝ÜĶ»Ęµø┐µŹóõĖ║µé©ńÜäÕ«×ķÖģÕÅśÕ╝éµĢ░µŹ«’╝ü
variants_df <-read.delim(
  file = '/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/CaTPS16_gviz_variants.tsv',
  header = T,
  sep = '\t'
  )

variants_df$chromosome[variants_df$chromosome == 'CanChrA20'] <- 'chr20'


# Õ”éµ×£ÕĀåÕÅĀõ╗Źµ£ēķŚ«ķóś’╝īĶĆāĶÖæÕłåĶĮ©ķüōµśŠńż║
snp_track <- AnnotationTrack(
  variants_df[variants_df$type == "SNP", ],
  name = "SNP",
  fill = "green", 
  lwd = 0,
  stacking = "squish"
)
all_tracks <- append(all_tracks, snp_track)

ins_track <- AnnotationTrack(
  variants_df[variants_df$type == "INS", ],
  name = "INS",
  fill = "red",
  lwd = 0,
  stacking = "squish"
)
all_tracks <- append(all_tracks, ins_track)
del_track <- AnnotationTrack(
  variants_df[variants_df$type == "DEL", ],
  name = "DEL", 
  fill = "blue",
  lwd = 0,
  stacking = "squish"
)
all_tracks <- append(all_tracks, del_track)


# ====================================
# 3. ń╗śÕłČÕøŠÕĮó 
# ====================================
print(length(all_tracks))  # µŻĆµ¤źĶĮ©Ķ┐╣ÕłŚĶĪ©µś»ÕÉ”µŁŻńĪ«
pdf("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/Hap_CaTPS16.pdf",
    width = 3.5,
    height = 2.5)

plot_title <- paste0("Chr20: ", start_coord, " - ", end_coord)

plotTracks(
  all_tracks,
  chromosome = target_chrom,
  from = start_coord,
  to = end_coord,
  main = plot_title,
  sizes = c(0.5, 1, 1, 1,1,1), 
  cex.main = 0.7,
  cex.axis = 0.6
)

dev.off()

```
###CaCYY28 on Chr11
```{r}
target_chrom <- "chr11"
start_coord <- 13524744
end_coord <- 13530703
chrom_length <- 14033347
custom_genome_name <- "Cangulatus" 

all_tracks <- list()

chrom_seqinfo <- Seqinfo(
  seqnames = target_chrom,
  seqlengths = chrom_length,
  genome = custom_genome_name 
)

# --- Track #1: IdeogramTrack (Õģ©ķĢ┐µ¤ōĶē▓õĮōÕøŠ) ---
full_chrom_band <- data.frame(
  chrom = target_chrom,            
  chromStart = 1,        
  chromEnd = chrom_length,            
  name = "chr11_full",
  gieStain = "gneg" 
)

ideo_track <- IdeogramTrack(
  genome = custom_genome_name, 
  chromosome = target_chrom,
  name = " ",
  bands = full_chrom_band,
  seqinfo = chrom_seqinfo, 
  outline = TRUE,
  fontcolor = "black",
  cex = 0.8,
  showTitle = FALSE,
  showId = FALSE
)
all_tracks <- append(all_tracks, ideo_track) 

# --- Track #2: GenomeAxisTrack (ÕØÉµĀćĶĮ┤) ---
axis_track <- GenomeAxisTrack(
  name = "Genomic Position (bp)",
  add53 = TRUE, 
  col = "black",
  cex = 0.8
)
all_tracks <- append(all_tracks, axis_track) 

# --- Track #3: GeneRegionTrack (Õ¤║ÕøĀń╗ōµ×ä - Hap1) ---
# µ©Īµŗ¤µĢ░µŹ«’╝ÜĶ»Ęµø┐µŹóõĖ║µé©ńÜäÕ«×ķÖģÕ¤║ÕøĀµ│©ķćŖµĢ░µŹ«’╝ü
gene_annot_df <- data.frame(
  chromosome = target_chrom,
  start = c(13526744, 13527762), 
  end = c(13527367, 13528703),   
  strand = "-", 
  feature = c("exon", "exon"),
  gene = "CaCYP28",
  transcript = "CaCYP28.T1",
  Haplotype = "Hap1" 
)

gene_track_hap1 <- GeneRegionTrack(
  gene_annot_df,
  chromosome = target_chrom,
  genome = custom_genome_name,
  seqinfo = chrom_seqinfo, 
  name = " ",
  fill = "lightblue",
  col = NULL,
  stacking = "squish", 
  featureAnnotation = "gene"
)
all_tracks <- append(all_tracks, gene_track_hap1) 

# --- Track #4: AnnotationTrack (ÕÅśÕ╝éõĮŹńé╣) ---
# µ©Īµŗ¤µĢ░µŹ«’╝ÜĶ»Ęµø┐µŹóõĖ║µé©ńÜäÕ«×ķÖģÕÅśÕ╝éµĢ░µŹ«’╝ü
variants_df <-read.delim(
  file = '/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/CaCYP28_gviz_variants.tsv',
  header = T,
  sep = '\t'
  )

variants_df$chromosome[variants_df$chromosome == 'CanChrA11'] <- 'chr11'

ins_track <- AnnotationTrack(
  variants_df[variants_df$type == "INS", ],
  name = "INS",
  fill = "red",
  lwd = 0,
  stacking = "squish"
)
all_tracks <- append(all_tracks, ins_track)
# ====================================
# 3. ń╗śÕłČÕøŠÕĮó 
# ====================================
print(length(all_tracks))  # µŻĆµ¤źĶĮ©Ķ┐╣ÕłŚĶĪ©µś»ÕÉ”µŁŻńĪ«
pdf("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/Hap_CaCYP28.pdf",
    width = 3.5,
    height = 2.5)

plot_title <- paste0("Chr11: ", start_coord, " - ", end_coord)

plotTracks(
  all_tracks,
  chromosome = target_chrom,
  from = start_coord,
  to = end_coord,
  main = plot_title,
  sizes = c(0.5, 1, 1,1), 
  cex.main = 0.7,
  cex.axis = 0.6
)

dev.off()

```

```{r}
library(tidyverse)
library(patchwork)
library(cowplot)

syri_dat <- read_tsv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/syri.out', col_names=FALSE)

colnames(syri_dat) = c('ref_chr','ref_start','ref_end','ref_strand',
                       'qry_strand','qry_chr','qry_start','qry_end',
                        'U_ID','P_ID','type','copy')

syri_dat <- syri_dat %>%
  mutate(ref_start = as.numeric(ref_start),
         ref_end = as.numeric(ref_end),
         qry_start = as.numeric(qry_start),
         qry_end = as.numeric(qry_end)
  )

table(syri_dat$type)

fai_dat <- read_tsv('/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/Celastrus_angulatus_A_genome.fasta.fai', 
                       col_names = c("chr", "length", "offset", "linebases", "linewidth"))
chr_lengths <- setNames(fai_dat$length, fai_dat$chr)

window_size <- 100000
target_chromosomes <- fai_dat$chr[1:3]  # ķĆēµŗ®ÕēŹ3õĖ¬µ¤ōĶē▓õĮōńö©õ║Äµ╝öńż║

# ÕćĮµĢ░’╝Üõ╗ÄsyriµĢ░µŹ«Ķ«Īń«ŚÕ»åÕ║”
calculate_variant_density <- function(syri_df, variant_type, chr_name, chr_length, window_size) {
  # ńŁøķĆēńē╣Õ«Üń▒╗Õ×ŗńÜäÕÅśÕ╝é
  if(variant_type == "SNP") {
    variant_data <- syri_df %>% filter(type %in% c("SNP", "SYN"))
    positions <- variant_data$ref_start[variant_data$ref_chr == chr_name]
  } else if(variant_type == "SV") {
    variant_data <- syri_df %>% filter(type %in% c("INV", "TRANS", "DUP", "NOTAL"))
    positions <- variant_data$ref_start[variant_data$ref_chr == chr_name]
  } else if(variant_type == "INV") {
    variant_data <- syri_df %>% filter(type == "INV")
    positions <- variant_data$ref_start[variant_data$ref_chr == chr_name]
  }
  
  if(length(positions) == 0) {
    return(data.frame(position = numeric(), density = numeric()))
  }
  
  windows <- seq(1, chr_length, by = window_size)
  densities <- sapply(windows, function(w) {
    sum(positions >= w & positions < (w + window_size))
  })
  
  data.frame(
    chr = chr_name,
    position = windows,
    density = densities,
    variant_type = variant_type
  )
}

# õĖ║µ»ÅõĖ¬µ¤ōĶē▓õĮōńö¤µłÉń╗äÕÉłÕøŠÕĮó
for(chr in target_chromosomes) {
  cat("Processing chromosome:", chr, "\n")
  
  chr_length <- chr_lengths[chr]
  if(is.na(chr_length)) next
  
  # Ķ«Īń«ŚÕÉäń¦ŹÕ»åÕ║”
  snp_density <- calculate_variant_density(syri_dat, "SNP", chr, chr_length, window_size)
  sv_density <- calculate_variant_density(syri_dat, "SV", chr, chr_length, window_size)
  inv_density <- calculate_variant_density(syri_dat, "INV", chr, chr_length, window_size)
  
  # ĶÄĘÕÅ¢Ķ»źµ¤ōĶē▓õĮōńÜäń╗ōµ×äÕÅśÕ╝é
  chr_sv <- syri_dat %>% 
    filter(ref_chr == chr, type %in% c("INV", "TRANS", "DUP"))
  
  # ÕłøÕ╗║Õ»åÕ║”ÕøŠ
  p_density <- ggplot() +
    # SNPÕ»åÕ║”
    geom_area(data = snp_density, aes(x = position, y = density, fill = "SNP"), alpha = 0.6) +
    # SVÕ»åÕ║”
    geom_area(data = sv_density, aes(x = position, y = density, fill = "SV"), alpha = 0.6) +
    scale_fill_manual(values = c("SNP" = "red", "SV" = "blue")) +
    labs(title = paste("Variant Density -", chr), y = "Density", fill = "Variant Type") +
    theme_minimal() +
    scale_x_continuous(limits = c(1, chr_length)) +
    theme(axis.title.x = element_blank(),
          legend.position = "top")
  
  # ÕłøÕ╗║ń╗ōµ×äÕÅśÕ╝éÕøŠ’╝łµ©Īõ╗┐plotsrķŻÄµĀ╝’╝ē
  p_sv <- ggplot() +
    # ĶāīµÖ»ńĮæµĀ╝
    geom_hline(yintercept = 1:3, color = "grey90", size = 0.2) +
    # ÕĆÆõĮŹ’╝łµ®ÖĶē▓’╝ē
    geom_segment(data = chr_sv %>% filter(type == "INV"),
                 aes(x = ref_start, xend = ref_end, y = 1, yend = 1),
                 color = "orange", size = 2, lineend = "round") +
    # µśōõĮŹ’╝łń┤½Ķē▓’╝ē
    geom_segment(data = chr_sv %>% filter(type == "TRANS"),
                 aes(x = ref_start, xend = ref_end, y = 2, yend = 2),
                 color = "purple", size = 2, lineend = "round") +
    # ķćŹÕżŹ’╝łń╗┐Ķē▓’╝ē
    geom_segment(data = chr_sv %>% filter(type == "DUP"),
                 aes(x = ref_start, xend = ref_end, y = 3, yend = 3),
                 color = "green", size = 2, lineend = "round") +
    # µĘ╗ÕŖĀµĀćńŁŠ
    geom_text(data = data.frame(y = 1:3, label = c("INV", "TRANS", "DUP"), 
                               x = rep(100000, 3)),
              aes(x = x, y = y, label = label), hjust = 0, size = 3) +
    labs(title = "Structural Variants", x = "Genomic Position (bp)", y = "") +
    theme_minimal() +
    scale_x_continuous(limits = c(1, chr_length)) +
    scale_y_continuous(breaks = 1:3, labels = c("Inversion", "Translocation", "Duplication")) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
  
  # ń╗äÕÉłÕøŠÕĮó’╝łń▒╗õ╝╝µ¢ćńī«Figure 5AÕĖāÕ▒Ć’╝ē
  combined_plot <- p_density / p_sv + 
    plot_layout(heights = c(2, 1)) +
    plot_annotation(
      title = paste("Chromosome", chr, "- Allelic Variations"),
      subtitle = "Integrated view of SNP density and structural variants",
      theme = theme(plot.title = element_text(size = 14, face = "bold"))
    )
  
  # õ┐ØÕŁśÕøŠÕĮó
  ggsave(paste0("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/figure5a_style_", chr, ".pdf"), combined_plot, 
         width = 14, height = 10, dpi = 300)
  ggsave(paste0("/Users/andy_zhu/Downloads/01-Recent_Tasks/01-AGIS/02-Cangulatus/03-Genome_Basic_Analysis/02-SV/figure5a_style_", chr, ".png"), combined_plot, 
         width = 14, height = 10, dpi = 300)
}

# ńö¤µłÉµ▒ćµĆ╗µŖźÕæŖ
summary_stats <- syri_dat %>%
  group_by(ref_chr, type) %>%
  summarise(count = n(), .groups = 'drop')

write_tsv(summary_stats, "variant_summary_stats.tsv")

cat("=== ń╗¤õĖĆÕØÉµĀćÕøŠÕĮóńö¤µłÉÕ«īµłÉ ===\n")
cat("ńö¤µłÉńÜäµ¢ćõ╗Č:\n")
cat("- figure5a_style_*.pdf/png: ÕÉäµ¤ōĶē▓õĮōńÜäń╗äÕÉłÕøŠÕĮó\n")
cat("- variant_summary_stats.tsv: ÕÅśÕ╝éń╗¤Ķ«ĪµæśĶ”ü\n")
cat("- ", prefix, "_plotsr.pdf: ÕÄ¤Õ¦ŗplotsrÕøŠÕĮó\n")


```


```{r}
library(plotsr)
```


